---
title: "Bioheart Deep Learning Manual Gating Tree"
author: "Elijah WIllie"
date: "2024-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = TRUE) 
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
```


```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(reticulate)
  library(keras)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(SKM)
  library(caret)
  library(ROCR)
  library(cvAUC)
  library(spicyR)
  library(cyCombine)
  library(ClassifyR)
  library(cowplot)
  library(psych)
  library(igraph)
  library(tensorflow)
  library(Spectrum)
  library(glmnet)
  library(ipflasso)
  library(prioritylasso)
  library(hopach)
  library(rpart)
  library(rpart.plot)
  library(data.tree)
  library(ggtree)
  library(ape)
  library(grpregOverlap)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
})

k = backend()     # this where the software used to break
source('~/Documents/PhD/bioheart_analysis/scripts/autoClass.R')
source('~/Documents/PhD/bioheart_analysis/scripts/helperFuncs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/deep_learning_helper_funcs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/customCV.R')
source('~/Documents/PhD/bioheart_analysis/scripts/predClusters.R')
source('~/Documents/PhD/bioheart_analysis/scripts/trainClassifier.R')
# source('~/Documents/PhD/DeepLearning_CyTOF/robust_ae.R')
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Read in the dataset
## Study 4
```{r}
# study 4
set.seed(1994)
df <- fread('../raw_data/bioheart_ct_cytof_data_b4_mg.csv',
              nThread = 7) %>%
  as.data.frame() 

# %>%
#   group_by(sample_id) %>%
#   slice_sample(n = 5000) 

# df <- df[which(df$mg_cell_type_distinct != ""), ]
# df <- df[which(df$mg_cell_type_clean != "/remove 34+ contam"), ]
# df <- df[which(df$mg_cell_type_clean != "CD3+ dirty"), ]
# df <- df[which(df$mg_cell_type_clean != "nonT"), ]
# df <- df[which(df$mg_cell_type_clean != "nonTnonB"), ]
# df <- df[which(df$mg_cell_type_distinct != "other"), ]

df[, useMarkers] <- log(abs(df[, useMarkers]) + 1)

# 
preProcValues <- preProcess(df[, useMarkers], method = c("range"))
# 
df[, useMarkers] <- predict(preProcValues, df[, useMarkers])

```

## Study 3
```{r}
df_3 <- fread('../raw_data/study_3_sub.csv',
              nThread = 7) %>% 
  as.data.frame() 

# %>%
#   group_by(sample_id)

df_3[, useMarkers] <- log(abs(df_3[, useMarkers]) + 1) 

df_3[, useMarkers] <- predict(preProcValues, df_3[, useMarkers])
```

# Get the reference sample for training
```{r}
res_ind <- ComputeReferenceSample(df, useMarkers)
```

# Get the encoding dimensions and the number of clusters 
```{r}
train_dat <- df[which(df$sample_id == res_ind$refSampleInd), useMarkers]
encoding_dim <- FuseSOM::computeGridSize(train_dat[, useMarkers])
```


# Define autoencoder model
```{r}
models <- autoEncoderModel(train_dat = train_dat, input_size = length(useMarkers),
                           encoding_dim = encoding_dim*2, l2_reg = 5E-04)

# model  <- robust_ae(train_dat[, useMarkers], encoding_dim = encoding_dim*2, batch_size = 16, epochs = 100)
```

# Predict on study 4
```{r}
df_norm <- predict(models$autoencoder_model, as.matrix(df[, useMarkers])) %>%
  as.data.frame()

colnames(df_norm) <- useMarkers

# df_latent <- predict(models$encoder_model, as.matrix(df[, useMarkers])) %>%
#   as.data.frame()
```

# Create SCE object for study 4 and cluster it
```{r}
# df_norm <- df_norm %>%
#   as.data.frame()
sce_norm <- SingleCellExperiment(assays = list(norm = t(df_norm[, useMarkers]),
                                               raw = t(df[, useMarkers])
),
colData = df %>% dplyr::select(-useMarkers))
# sce_norm$clusters <- y_pred

nclust <- length(unique(sce_norm$mg_cell_type_clean))
```

# Predict on study 3
```{r}
df_3_norm <- predict(models$autoencoder_model, as.matrix(df_3[, useMarkers])) %>%
  as.data.frame()
colnames(df_3_norm) <- useMarkers

# df_3_latent <- predict(models$encoder_model, as.matrix(df_3[, useMarkers])) %>%
#   as.data.frame()
```
# Train clustering model
## Select training data 
```{r}
set.seed(1994)
# df_norm <- df_norm
# df_norm$clusters_mg <- factor(sce_norm$mg_cell_type_distinct)
df_norm$clusters <- factor(sce_norm$mg_cell_type_distinct)
df_norm$sample_id <- sce_norm$sample_id
# df_norm <- df_norm[df_norm$clusters != "", ]
# df_norm <- df_norm[df_norm$clusters != "other", ]
df_norm$clusters <- droplevels(df_norm$clusters)

df_norm_sub <- df_norm[which(df$sample_id == res_ind$refSampleInd), ]

# min_val <- min(table(df_norm$clusters))
# 
# df_norm_sub <- df_norm %>%
#   group_by(clusters) %>%
#   slice_sample(n = min_val, replace = F)
```


```{r}
y = onehot_encode(as.numeric(droplevels(df_norm_sub$clusters)) - 1)  


# perform a fit and predict [ elmNNRcpp ]
#----------------------------------------

fit_elm_class = elm_train(as.matrix(df_norm_sub[, useMarkers]), y, nhid = 10000, actfun = 'tansig',
                          
                          init_weights = "normal_gaussian", bias = FALSE, verbose = TRUE)

df_3_clusters = elm_predict(fit_elm_class, as.matrix(df_3_norm), normalize = FALSE)

df_3_clusters = as.factor(max.col(df_3_clusters, ties.method = "random"))

levels(df_3_clusters) <- droplevels(df_norm_sub$clusters) %>% unique()
table(df_3_clusters)
```


## Fit SKM neural net model
```{r}
nclust <- length(unique(df_norm_sub$clusters))
nnet_model <- SKM::deep_learning(
  y = df_norm_sub$clusters,
  x =  as.matrix(df_norm_sub[, useMarkers]),
  epochs_number = 100,
  layers = list(list(neurons_number = 12, 
                      neurons_proportion = NULL, 
                       activation = "softplus",
                     ridge_penalty = 5E-04),
                list(neurons_number = 6, 
                      neurons_proportion = NULL, 
                       activation = "softplus",
                     ridge_penalty = 5E-04),
                list(neurons_number = 3, 
                      neurons_proportion = NULL, 
                       activation = "softplus",
                     ridge_penalty = 5E-04)),
  optimizer = 'adamax', seed = 1994, 
  loss_function = 'categorical_crossentropy', 
  batch_size = 16, learning_rate = 0.005,
  verbose = TRUE
)
```

## Now preedict on study 3
```{r}
df_3_clusters <- nnet_model %>%
    predict(as.matrix(df_3_norm))
df_3_clusters <- df_3_clusters$predicted
```

```{r}
study_4 <- table(df$mg_cell_type_distinct)/sum(table(df$mg_cell_type_distinct))
study_4 <- table(df$mg_cell_type_distinct)/sum(table(df$mg_cell_type_distinct))
study_3 <- table(df_3_clusters)/sum(table(df_3_clusters))
cbind(study_4, study_3)
```

## Compute features on study 3
### Generate SCE object and populate clusters
```{r}
sce_3_norm <- SingleCellExperiment(assays = list(raw = t(df_3[,useMarkers]),
                                                 norm = t(df_3_norm[, useMarkers])
                                          ),
                                 colData = df_3[, 1:10])

sce_3_norm$clusters <- df_3_clusters
```

# Compute training features
## Proportions
```{r}
sce_norm_sub <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
sce_norm_sub <- sce_norm_sub[, sce_norm_sub$mg_cell_type_distinct != ""]
# do prediction
data <- getProp(sce_norm_sub,
                feature = "mg_cell_type_distinct", imageID = "sample_id", logit = F)

dat_prop <- data
data <- 2*asin(sqrt(data))

clinicaldata <- colData(sce_norm_sub) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

row_names <- rownames(data)

Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"]
GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")]

clinicaldata_info <- cbind(Age, GenderM)

condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])


data_logit <- getProp(sce_norm_sub,
                feature = "mg_cell_type_distinct", imageID = "sample_id", logit = T)
colnames(data_logit) <- paste0(colnames(data_logit), "_logit")
colnames(data) <- paste0(colnames(data), "_prop")
colnames(dat_prop) <- colnames(data)
```

## Mean expression
```{r}
t <- assay(sce_norm_sub, 'norm') %>%
  t() %>%
  as.data.frame()
t$sample_id <- sce_norm_sub$sample_id

data_mean <- t %>% 
  dplyr::group_by(sample_id) %>%
  dplyr::summarise_at(vars(useMarkers), mean) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

data_median <- t %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), median) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

rownames(data_mean) <- clinicaldata$sample_id
rownames(data_median) <- clinicaldata$sample_id
colnames(data_mean) <- paste0(colnames(data_mean), "_mean")
colnames(data_median) <- paste0(colnames(data_median), "_median")
```

## Generate features for Study 3
### Proportions
```{r}
sce_3_norm_sub <- sce_3_norm[, sce_3_norm$clusters != "other"]
sce_3_norm_sub <- sce_3_norm_sub[, sce_3_norm_sub$clusters != ""]
sce_3_norm_sub$clusters <- droplevels(sce_3_norm_sub$clusters)

data_3 <- getProp(sce_3_norm_sub,
                feature = "clusters", imageID = "sample_id", logit = F)


data_3 <- 2*asin(sqrt(data_3))
# 
# data_3_2 <- getProp(sce_3_norm,
#                   feature = "clusters", imageID = "sample_id", logit = T)
# data_3 <- cbind(data_3, data_3_2)
colnames(data_3) <- janitor::make_clean_names(colnames(data_3))

clinicaldata_3 <- colData(sce_3_norm_sub) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Age, Gender) %>%
  distinct()
clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)

row_names_3 <- rownames(data_3)

Age <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)

data_3_logit <- getProp(sce_3_norm_sub,
                feature = "clusters", imageID = "sample_id", logit = T)

condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])

colnames(data_3_logit) <- paste0(colnames(data_3_logit), "_logit")
colnames(data_3) <- paste0(colnames(data_3), "_prop")
```


### Mean expression
```{r}
t_3 <- assay(sce_3_norm_sub, 'norm') %>%
  t() %>%
  as.data.frame()
t_3$sample_id <- sce_3_norm_sub$sample_id

t_3_mean <- t_3 %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), mean) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

t_3_median <- t_3 %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), median) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

rownames(t_3_mean) <- clinicaldata_3$sample_id
rownames(t_3_median) <- clinicaldata_3$sample_id

colnames(t_3_mean) <- paste0(colnames(t_3_mean), "_mean")
colnames(t_3_median) <- paste0(colnames(t_3_median), "_median")
```

# Generate the MG tree
```{r, eval=TRUE}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_prop')
    `14+ monos` <- myeloid$AddChild('14+ monos_prop')
    `16+ monos` <- myeloid$AddChild('16+ monos_prop') 
  DCs <- root$AddChild('DCs_prop')
    pDCs <- DCs$AddChild('pDCs_prop')
    mDCs <- DCs$AddChild('mDCs_prop')
      `CD141+ DCs` <- mDCs$AddChild('CD141+ DCs_prop')
      `CD1c+ DCs` <- mDCs$AddChild('CD1c+ DCs_prop')
  NK <- root$AddChild('NK_prop')
  `B cells` <- root$AddChild('B cells_prop')
  CD3 <- root$AddChild('CD3_prop')
    CD8hi <- CD3$AddChild('CD8hi_prop')
    CD8lo <- CD3$AddChild('CD8lo_prop')
    CD4 <- CD3$AddChild('CD4_prop')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_prop')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_prop')
  # other <- root$AddChild('other_prop')
```

```{r}
hc_prop <- treekoR:::findChildren(ggtree(as.phylo(root), 
                                           ladderize = F, layout = 'dendrogram'))
# Extract clusters and assign group numbers
subs_prop <- hc_prop$data$clusters
prop_dat <- hc_prop$data
prop_dat$group <- 1:length(subs_prop)
```

```{r, eval=TRUE}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_logit')
    `14+ monos` <- myeloid$AddChild('14+ monos_logit')
    `16+ monos` <- myeloid$AddChild('16+ monos_logit') 
  DCs <- root$AddChild('DCs_logit')
    pDCs <- DCs$AddChild('pDCs_logit')
    mDCs <- DCs$AddChild('mDCs_logit')
      `CD141+ DCs` <- mDCs$AddChild('CD141+ DCs_logit')
      `CD1c+ DCs` <- mDCs$AddChild('CD1c+ DCs_logit')
  NK <- root$AddChild('NK_logit')
  `B cells` <- root$AddChild('B cells_logit')
  CD3 <- root$AddChild('CD3_logit')
    CD8hi <- CD3$AddChild('CD8hi_logit')
    CD8lo <- CD3$AddChild('CD8lo_logit')
    CD4 <- CD3$AddChild('CD4_logit')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_logit')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_logit')
  # other <- root$AddChild('other_logit')
```

```{r}
hc_logit <- treekoR:::findChildren(ggtree(as.phylo(root), 
                                           ladderize = F, layout = 'dendrogram'))
# Extract clusters and assign group numbers
subs_logit <- hc_logit$data$clusters
logit_dat <- hc_logit$data
logit_dat$group <- 1:length(subs_logit)
```

```{r}
# Compute pairwise Euclidean distances for gene means per cell type data
d_mean <- parallelDist::parallelDist(t(data_mean), method = 'euclidean')
# Hierarchical clustering of gene means per cell type data
hc_mean <- hclust(d_mean, method = 'average')
# Extract clusters from the dendrogram
hc_mean <- treekoR:::findChildren(ggtree(as.phylo(hc_mean),
                                           ladderize = F, layout = 'dendrogram'))
# Extract clusters and assign group numbers
subs_mean <- hc_mean$data$clusters
mean_dat <- hc_mean$data
mean_dat$group <- 1:length(subs_mean)
```

```{r}
nclust <- ncol(data)
# Prepare variable groups for overlapping group Lasso
var_groups<- list()
n <- length(subs_logit)
# n2 <- 2*n
for(i in 1:length(subs_logit)){
    var_groups[[i]] <- which(make_clean_names(colnames(data_logit)) %in% 
                               make_clean_names(subs_logit[[i]]))
    # var_groups[[i+n]] <- which(make_clean_names(colnames(data_logit)) %in% 
    #                              make_clean_names(subs_logit[[i]])) + nclust
  }
  
for(i in 1:length(subs_mean)){
    var_groups[[i+n]] <- which(colnames(data_mean) %in% subs_mean[[i]]) + nclust
  }
```

```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, data_mean) 
colnames(X_train) <- janitor::make_clean_names(colnames(X_train))
  
  # Scale the combined data
scaleVals <- preProcess(X_train, method = c('range'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
  
  # Extract response variable
y_train <- condition %>%
    as.numeric()
```

```{r}
# Cross-validation for overlapping group Lasso
alpha <- 1
cvfit <- cv.grpregOverlap(X_train, y_train, var_groups, penalty = 'grLasso', 
                            family = 'binomial',  alpha = alpha, nfolds = 10,
                            seed = 999)
# Plot cross-validation results
plot(cvfit)
  
# Fit overlapping group Lasso model
fit <- grpregOverlap(X_train, y_train, var_groups,
                       family='binomial', alpha = alpha,
                       returnX.latent = T, returnOverlap = FALSE, 
                       lambda = cvfit$lambda.min, penalty = 'grLasso')
```

## Set up test data
```{r}
X_test_study_3 <- cbind(data_3_logit, t_3_mean) %>%
  as.matrix()
colnames(X_test_study_3) <- janitor::make_clean_names(colnames(X_test_study_3))
X_test_study_3 <- predict(scaleVals, X_test_study_3)
```

## Predict on test data
```{r}
pred_study_3 <- predict(fit, X = X_test_study_3,
                        type = "response")

write.csv(as.data.frame(pred_study_3),
          '../auc_data/mg_model_auc_study_3.csv')

pred_study_3 <- prediction(pred_study_3, condition_3)
pred_study_3_perf <- ROCR::performance(pred_study_3, "tpr", "fpr")


plt_dat_3 = data.frame(
  FPR = pred_study_3_perf@x.values[[1]],
  TPR = pred_study_3_perf@y.values[[1]]
)

auc_perf_study_3 <- ROCR::performance(pred_study_3, measure = "auc")
auc_study_3 <- auc_perf_study_3@y.values[[1]]
```

```{r}
# Generate AUC plot
ggplot(plt_dat_3, aes(x = FPR, y = TPR)) +
  geom_line(colour = "blue") +
  labs(x = pred_study_3_perf@x.name, y = pred_study_3_perf@y.name) +
  geom_abline(slope = 1, intercept = 0) + theme_bw() +
  theme(
    plot.title = element_text(color="Black", size=16, face="bold", hjust = 0.5),
    plot.subtitle = element_text(color = "red", size = 16, hjust = 0.5),
    axis.title.x = element_text(color="Black", size=16),
    axis.title.y = element_text(color="Black", size=16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 16),
    strip.text.x = element_text(size = 16),
    legend.title = element_text(size=16), #change legend title font size
    legend.text = element_text(size=16) #change legend text font size)
  )  + ggtitle(paste("Overlap Group Lasso MG- Study 3 - AUC =", round(auc_study_3, 2)))
```

# Plot overlap tree
```{r}
coefs <- coef(fit) %>%
  as.matrix() %>%
  as.data.frame()
coefs$feature <- rownames(coefs)
res_mp <- MP_gLasso(fit = fit, group = fit$grp.vec, lambda.type = "min", sort.type = "max")
```


## Proportions
```{r, fig.height=6, fig.width=12}
p <- plotOverlapTree(hc = hc_logit, coefs = coefs, type = 'logit',
                     data = res_mp$data, heatmap = F)
colnames(dat_prop) <- str_replace(colnames(dat_prop), '_prop', '')
```


```{r, fig.height=9, fig.width=10}
# generate the heatmap matrix
hmat <- generateHeatmapMatrix(t, markers = useMarkers, 
                              clusters = sce_norm_sub$mg_cell_type_distinct, threshold = 2) %>%
  as.data.frame()
rownames(hmat) <- janitor::make_clean_names(rownames(hmat))
colnames(hmat) <- janitor::make_clean_names(colnames(hmat))

pdf(file = "../Plots/Manuscript_Plots/mg_tree.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 10) # The height of the plot in inches

gheatmap(p, data = hmat, offset = 0.05, hjust = 0.75, colnames_offset_y = 0, 
         color = 'white',legend_title="Proportions") +
  # vexpand(.1, -1) + 
      theme(legend.title = element_text(color = "black", size = 10),
            legend.text = element_text(color = "black", size = 10)) + 
  scale_fill_gradientn(colours = eval(formals(pheatmap::pheatmap)$color, 
                                      envir=environment(pheatmap::pheatmap)), 
                       name = "Mean Expression")
dev.off()
```


## Means
```{r, fig.height=6, fig.width=12}
p.mean <- plotOverlapTree(hc = hc_mean, coefs = coefs, type = 'mean',
                     data = res_mp$data, heatmap = F,
                     offset <- length(hc_logit$data$clusters))
```

```{r}
gheatmap(p.mean, data = t(hmat), offset = 0.05, hjust = 0.75, colnames_offset_y = 0, 
         color = 'white',legend_title="Proportions") +
  # vexpand(.1, -1) + 
      theme(legend.title = element_text(color = "black", size = 10),
            legend.text = element_text(color = "black", size = 10)) + 
  scale_fill_gradientn(colours = eval(formals(pheatmap::pheatmap)$color, 
                                      envir=environment(pheatmap::pheatmap)), 
                       name = "Mean Expression")
```

# Fit a baseline Lasso model
```{r}
set.seed(999)
cv.lasso <- cv.glmnet(X_train, condition, family='binomial', alpha=1, parallel=TRUE, 
                      standardize=F, relax = F, type.measure = 'auc',nfolds = 10)
plot(cv.lasso)
coef(cv.lasso, s='lambda.min')
coef <- coef(cv.lasso, s='lambda.min')
selected_attributes <- (coef@i[-1]+1)

```

```{r}
pred_test <- predict(cv.lasso, newx = X_test_study_3, 
                     type = "response", s = 'lambda.min')

write.csv(as.data.frame(pred_test),
          '../auc_data/mg_model_baseline_auc_study_3.csv')


pred_fs <- prediction(pred_test, condition_3)
perf_fs <- ROCR::performance(pred_fs, "tpr", "fpr")


plt_dat = data.frame(
  FPR = perf_fs@x.values[[1]],
  TPR = perf_fs@y.values[[1]]
)

auc_perf <- ROCR::performance(pred_fs, measure = "auc")
auc <- auc_perf@y.values[[1]]

# generate AUC plot
p_auc <- ggplot(plt_dat, aes(x = FPR, y = TPR)) +
  geom_line(colour = "blue") +
  labs(x = perf_fs@x.name, y = perf_fs@y.name) +
  geom_abline(slope = 1, intercept = 0) + theme_bw() +
  theme(
    plot.title = element_text(color="Black", size=16, face="bold", hjust = 0.5),
    plot.subtitle = element_text(color = "red", size = 16, hjust = 0.5),
    axis.title.x = element_text(color="Black", size=16),
    axis.title.y = element_text(color="Black", size=16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 16),
    strip.text.x = element_text(size = 16),
    legend.title = element_text(size=16), #change legend title font size
    legend.text = element_text(size=16) #change legend text font size)
  )  + ggtitle(paste("Test Set - AUC =", round(auc, 2)))

p_auc
```

