---
title: "Bioheart Prediction Manual Gating"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(keras3)
  library(reticulate)
  use_virtualenv('r-keras', required = TRUE)
  library(tensorflow)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(caret)
  library(ROCR)
  library(psych)
  library(tensorflow)
  library(glmnet)
  library(data.tree)
  library(ggtree)
  library(ape)
  library(grpregOverlap)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
})
source('~/Documents/PhD/bioheart_analysis/scripts/helperFuncs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/customCV.R')
set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA-DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c-PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Load normalised data
```{r}
load('../sce_dat/study_4_MMD_VAE_sce.RData')
load('../sce_dat/study_3_MMD_VAE_sce.RData')
```

# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
```


```{r, fig.height=10, fig.width=16, eval=FALSE}
sce_norm_sub <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
sce_norm_sub <- sce_norm_sub[, sce_norm_sub$mg_cell_type_distinct != ""]
ggplotify::as.ggplot(scater::plotGroupedHeatmap(sce_norm_sub, 
                           features = useMarkers, 
                           group = "mg_cell_type_distinct", 
                           block = "mg_cell_type_distinct",
                           exprs_values = "norm",
                           center = TRUE, 
                           scale = TRUE, 
                           zlim = c(-2,2),
                           cluster_rows = FALSE)) + 
  # theme_bw() +
  # theme_minimal() +
  ggtitle(label = "Heatmap of normalised manually gated clusters expression") +
    theme(
          plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
    )
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- assay(sce_norm, "raw") %>%
  t() %>%
  as.data.frame() %>%
  dplyr::mutate(cellTypes = sce_norm$mg_cell_type_distinct)
  # mutate(sample_id = sce_norm$sample_id) %>%
  # group_by(clusters) %>%
  # dplyr::filter(sample_id %in%  c('586', '804')) %>%
  # dplyr::select(-sample_id)


test_x <- assay(sce_3_norm, type = "counts") %>%
  t() %>%
  as.data.frame()
```

## Fit SVM model
```{r, eval=T}
df_3_clusters <- cellTypeClassifier(train_x, test_x, model = 'lda')

table(df_3_clusters)

sce_3_norm$mg_clusters <- df_3_clusters
```

# Setup Clinical information
## Study 4
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

# row_names <- rownames(data)

# Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       "Age"]
# GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       c("GenderM")]

# clinicaldata_info <- cbind(Age, GenderM)
```

## Study 3
```{r}
clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
# clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)
clinicaldata_3$Gensini_bin <- as.factor(clinicaldata_3$Gensini_bin)
# row_names_3 <- rownames(data_3)

# Age_3 <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
#                                       "Age"]
# GenderM_3 <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
#                                       c("GenderM")]
# 
# clinicaldata_info_3 <- cbind(Age, GenderM)
```


# Compute training features
## Proportions
```{r}
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != ""]
# sce_norm$mg_cell_type_distinct <- droplevels(sce_norm$mg_cell_type_distinct)
data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', 
                              sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])

data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = F, useMarkers = useMarkers)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = "raw",
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

markerMeanCellType_rsc <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = 'norm_rescaled',
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```


## Compute features on study 3

## Proportions
```{r}
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != "other"]
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != ""]
sce_3_norm$mg_clusters <- droplevels(sce_3_norm$mg_clusters)

data_3_logit <- computeFeatures(sce = sce_3_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_3 <- rownames(data_3_logit)
condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])
```

## Means
```{r, eval=T}
markerMeanCellType_3 <- computeFeatures(sce = sce_3_norm, featureType = 'mean',  assay = "raw",
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

# Generate Tree
```{r, eval=T}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_logit')
    `14+ monos` <- myeloid$AddChild('14+ monos_logit')
    `16+ monos` <- myeloid$AddChild('16+ monos_logit') 
    pDCs <- myeloid$AddChild('pDCs_logit')
    `CD141+ DCs` <- myeloid$AddChild('CD141+ DCs_logit')
    `CD1c+ DCs` <- myeloid$AddChild('CD1c+ DCs_logit')
  NK <- root$AddChild('NK_logit')
  `B cells` <- root$AddChild('B cells_logit')
  CD3 <- root$AddChild('CD3_logit')
    CD8hi <- CD3$AddChild('CD8hi_logit')
    CD8lo <- CD3$AddChild('CD8lo_logit')
    CD4 <- CD3$AddChild('CD4_logit')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_logit')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_logit')
  # other <- root$AddChild('other_logit')
```

# Generate grouping structures
```{r}
nclust <- ncol(data_logit)
root <- as.phylo(root)
order <- root$tip.label
tree <- treekoR:::findChildren(ggtree(as.phylo(root),
                                            ladderize = F, layout = 'dendrogram'))

groups <- generateGroups(tree = tree, nclust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Add the clinical data in
```{r, eval=FALSE}
var_groups[[length(var_groups) + 1]] <- 309:310
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
  # Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()

# Extract response variable
y_train <- condition %>%
    as.numeric()
```

## Testing
```{r}
X_test_study_3 <- cbind(data_3_logit, markerMeanCellType_3) %>%
  as.matrix()
# colnames(X_test_study_3) <- janitor::make_clean_names(colnames(X_test_study_3))
X_test_study_3 <- predict(scaleVals, X_test_study_3)
```

## Fit Overlap model
```{r}
fit <- fitModel(x_train = X_train, y_train = y_train, groups = groups)
```

## Predict on the training data
```{r}
train_auc <- plotAUC(fit = fit$fit, x_test = X_train, y_test = condition, title = "Study 4 MG - All AUC =")
train_auc$plot
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, x_test = X_test_study_3, y_test = condition_3, title = "Study 3 MG - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../auc_data/mg_glasso_auc_all_study_3_baseline.csv')

```



# Visualise Resulting Tree
```{r, fig.height=16, fig.width=16}
p <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", training_data = X_train,
                   nodes_to_remove =  c("Root", "Myeloid", "CD3", "CD4"), title = "Baseline Manual Gating Feature Tree")

ggsave(file="../Plots/baseline_mg_tree.svg", plot=p, width=18, height=16, dpi = 1200)
```

## Study 4
```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType_rsc, clinicaldata = clinicaldata, 
                         outcome = "Gensini_bin")

if(sum(stats_mean$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_mean$stats, sigFeatures = stats_mean$sig_features, outcome = "Gensini bin")
}
```

```{r, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop <- getSigFeatures(fit$fit, type = 'prop', prop = data, 
                              clinicaldata = clinicaldata, 
                         outcome = "Gensini_bin")
if(sum(stats_prop$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_prop$stats, sigFeatures = stats_prop$sig_features, outcome = "Gensini bin")
}
```
## Study 3
```{r, eval=T, fig.height=10, fig.width=16, eval=TRUE}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean_3 <- getSigFeatures(fit, type = 'mean',  
                         mean = markerMeanCellType_3, clinicaldata = clinicaldata_3, 
                         outcome = "Gensini_bin")

if(sum(stats_mean_3$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_mean_3$stats, sigFeatures = stats_mean_3$sig_features, outcome = "Gensini bin")
}
```

```{r, eval=T, fig.height=10, fig.width=16, eval=FALSE}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop_3 <- getSigFeatures(fit, type = 'prop',  
                         prop = data_3_logit, clinicaldata = clinicaldata_3, 
                         outcome = "Gensini_bin")

if(sum(stats_prop_3$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_prop_3$stats, sigFeatures = stats_prop_3$sig_features, outcome = "Gensini bin")
}
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test_study_3 <- data_3_logit %>%
  as.matrix()
X_test_study_3 <- predict(scaleVals, X_test_study_3)
```

```{r}
groups <- generateGroups(tree = tree, nclust = nclust, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
fit <- fitModel(x_train = X_train, y_train = y_train, groups = groups)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, x_test = X_test_study_3, y_test = condition_3, title = "Study 3 MG - Proportions AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../auc_data/mg_glasso_auc_prop_study_3_baseline.csv')

```

## Look at tree plot
```{r, fig.height=10, fig.width=16}
visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", training_data = X_train,
                   nodes_to_remove =  c("Root", "Myeloid", "CD3", "CD4"), heatmap = FALSE)
```

```{r, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop <- getSigFeatures(fit$fit, type = 'prop', prop = data_logit, 
                              clinicaldata = clinicaldata, 
                         outcome = "Gensini_bin")

if(sum(stats_prop$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_prop$stats, sigFeatures = stats_prop$sig_features, outcome = "Gensini bin")
}

```

```{r, eval=T, fig.height=10, fig.width=16, eval=TRUE}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop_3 <- getSigFeatures(fit$fit, type = 'prop',  
                         prop = data_3_logit, clinicaldata = clinicaldata_3, 
                         outcome = "Gensini_bin")

if(sum(stats_prop_3$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_prop_3$stats, sigFeatures = stats_prop_3$sig_features, outcome = "Gensini bin")
}
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test_study_3 <- markerMeanCellType_3 %>%
  as.matrix()
X_test_study_3 <- predict(scaleVals, X_test_study_3)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(x_train = X_train, y_train = y_train, groups = groups)
```

## Predict on the training data
```{r}
train_auc <- plotAUC(fit = fit$fit, x_test = X_train, y_test = condition, title = "Study 4 MG - Means AUC =")
train_auc$plot
```
## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, x_test = X_test_study_3, y_test = condition_3, title = "Study 3 MG - Means AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../auc_data/mg_glasso_auc_means_study_3_baseline.csv')

```


```{r, fig.height=10, fig.width=16}
order <- gsub("_logit", "", order)
order <- gsub("_", " ", order)
plotHeatmap(fit$fit, type = "MG", order = order)
```

```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType, clinicaldata = clinicaldata, 
                         outcome = "Gensini_bin")

if(sum(stats_mean$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_mean$stats, sigFeatures = stats_mean$sig_features, outcome = "Gensini bin")
}
```

## Study 3
```{r, eval=T, fig.height=10, fig.width=16, eval=FALSE}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean_3 <- getSigFeatures(fit, type = 'mean',  
                         mean = markerMeanCellType_3, clinicaldata = clinicaldata_3, 
                         outcome = "Gensini_bin")

if(sum(stats_mean_3$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_mean_3$stats, sigFeatures = stats_mean_3$sig_features, outcome = "Gensini bin")
}
```