---
title: "Bioheart Deep Learning Prediction Manual Gating"
author: "Elijah WIllie"
date: "2024-03-15"
output: html_document
---

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(reticulate)
  library(keras)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(SKM)
  library(caret)
  library(ROCR)
  library(cvAUC)
  library(spicyR)
  library(cyCombine)
  library(ClassifyR)
  library(cowplot)
  library(psych)
  library(igraph)
  library(tensorflow)
  library(Spectrum)
  library(glmnet)
  library(ipflasso)
  library(prioritylasso)
  library(hopach)
  library(rpart)
  library(rpart.plot)
  # library(ruta)
  library(ggtree)
  library(ape)
  library(grpregOverlap)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
})

k = backend()     # this where the software used to break
source('~/Documents/PhD/bioheart_analysis/scripts/autoClass.R')
source('~/Documents/PhD/bioheart_analysis/scripts/helperFuncs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/deep_learning_helper_funcs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/customCV.R')
source('~/Documents/PhD/bioheart_analysis/scripts/predClusters.R')
source('~/Documents/PhD/bioheart_analysis/scripts/trainClassifier.R')
# source('~/Documents/PhD/DeepLearning_CyTOF/robust_ae.R')
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Load in the normalised datasets
```{r}
load('../sce_dat/study_4_autoEncoder_sce.RData')
load('../sce_dat/study_3_autoEncoder_sce.RData')
```



# Compute training features
## Proportions
```{r}
# do prediction
data <- getProp(sce_norm,
                feature = "mg_cell_type_distinct", imageID = "sample_id", logit = F)

dat_prop <- data
data <- 2*asin(sqrt(data))

clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

row_names <- rownames(data)

Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"]
GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")]

clinicaldata_info <- cbind(Age, GenderM)

condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])


data_logit <- getProp(sce_norm,
                feature = "mg_cell_type_distinct", imageID = "sample_id", logit = T)
colnames(data_logit) <- paste0(colnames(data_logit), "_logit")
colnames(data) <- paste0(colnames(data), "_prop")
colnames(dat_prop) <- colnames(data)
```

```{r}
df_norm <- assay(sce_norm, 'norm') %>%
  t() %>%
  as.data.frame()

df_raw <- assay(sce_norm, 'raw') %>%
  t() %>%
  as.data.frame()

df_3_norm <- assay(sce_3_norm, 'norm') %>%
  t() %>%
  as.data.frame()
```


## Mean expression
```{r}
t <- df_norm
t$sample_id <- sce_norm$sample_id

data_mean <- t %>% 
  dplyr::group_by(sample_id) %>%
  dplyr::summarise_at(vars(useMarkers), mean) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

data_median <- t %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), median) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

rownames(data_mean) <- clinicaldata$sample_id
rownames(data_median) <- clinicaldata$sample_id
colnames(data_mean) <- paste0(colnames(data_mean), "_mean")
colnames(data_median) <- paste0(colnames(data_median), "_median")
```

# Look at CV performance
```{r, eval=F}
res.cv <- KFoldCustomEnsemble(dat_prop = data, dat_logit = data_logit, 
                              dat_mean = data_mean, linkage = 'ward.D2',
                              train_y = condition, alpha = 0.5, 
                              nRepeats = 10, k = 10)
```

# Train clustering model
## Get the reference sample for training
```{r}
df_raw$sample_id <- sce_norm$sample_id
res_ind <- ComputeReferenceSample(df_raw, useMarkers)
```
## Select training data 
```{r}
set.seed(1994)
# df_norm <- df_norm
# df_norm$clusters_mg <- factor(sce_norm$mg_cell_type_distinct)
df_norm$clusters <- factor(sce_norm$mg_cell_type_distinct)
df_norm$sample_id <- sce_norm$sample_id

df_norm_sub <- df_norm[which(sce_norm$sample_id == res_ind$refSampleInd), ]

# min_val <- min(table(df_norm$clusters))
# 
# df_norm_sub <- df_norm %>%
#   group_by(clusters) %>%
#   slice_sample(n = 1000, replace = F)
```

## Fit SKM neural net model
```{r}
nclust <- length(unique(df_norm_sub$clusters))
nnet_model <- SKM::deep_learning(
  y = df_norm_sub$clusters,
  x =  as.matrix(df_norm_sub[, useMarkers]),
  epochs_number = 100,
  layers = list(list(neurons_number = 12, 
                      neurons_proportion = NULL, 
                       activation = "softplus"),
                list(neurons_number = 6, 
                      neurons_proportion = NULL, 
                       activation = "softplus"),
                list(neurons_number = 3, 
                      neurons_proportion = NULL, 
                       activation = "softplus"),
                list(neurons_number = nclust-1, 
                      neurons_proportion = NULL, 
                       activation = "softmax")),
  optimizer = 'adamax', seed = 1994, 
  loss_function = 'categorical_crossentropy', 
  batch_size = 16, learning_rate = 0.005,
  verbose = TRUE
)
```

## Now preedict on study 3
```{r}
df_3_clusters <- nnet_model %>%
    predict(as.matrix(df_3_norm))
df_3_clusters <- df_3_clusters$predicted
```

```{r}
study_4 <- table(sce_norm$mg_cell_type_distinct)/sum(table(sce_norm$mg_cell_type_distinct))
study_3 <- table(df_3_clusters)/sum(table(df_3_clusters))
cbind(study_4, study_3)
```


## Compute features on study 3
### Generate SCE object and populate clusters
```{r}
sce_3_norm$clusters_mg <- df_3_clusters
```

## Generate features
### Proportions
```{r}
data_3 <- getProp(sce_3_norm,
                feature = "clusters_mg", imageID = "sample_id", logit = F)


data_3 <- 2*asin(sqrt(data_3))
# 
# data_3_2 <- getProp(sce_3_norm,
#                   feature = "clusters", imageID = "sample_id", logit = T)
# data_3 <- cbind(data_3, data_3_2)
# colnames(data_3) <- janitor::make_clean_names(colnames(data_3))

clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Age, Gender) %>%
  distinct()
clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)

row_names_3 <- rownames(data_3)

Age_3 <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM_3 <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)

data_3_logit <- getProp(sce_3_norm,
                feature = "clusters_mg", imageID = "sample_id", logit = T)

condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])

colnames(data_3_logit) <- paste0(colnames(data_3_logit), "_logit")
colnames(data_3) <- paste0(colnames(data_3), "_prop")
```


### Mean expression
```{r}
t_3 <- df_3_norm
t_3$sample_id <- sce_3_norm$sample_id

t_3_mean <- t_3 %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), mean) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

t_3_median <- t_3 %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), median) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

rownames(t_3_mean) <- clinicaldata_3$sample_id
rownames(t_3_median) <- clinicaldata_3$sample_id

colnames(t_3_mean) <- paste0(colnames(t_3_mean), "_mean")
colnames(t_3_median) <- paste0(colnames(t_3_median), "_median")
```

# Overlap Lasso Modelling
## Fit model
```{r}
nclust <- length(unique(sce_3_norm$clusters_mg))
overlapModel <- runOverlapLasso(data, data_logit, data_mean,
                                nclust = nclust, alpha = 0.5,
                                linkage = 'average', clinical = F, 
                                clinicalData = cbind(Age, GenderM))
fit <- overlapModel$fit
scaleVals <- overlapModel$scaling
```

## Set up test data
```{r}
X_test_study_3 <- cbind(data_3_logit, t_3_mean, Age_3, GenderM_3)
# colnames(X_test_study_3) <- janitor::make_clean_names(colnames(X_test_study_3))
X_test_study_3[, 1:39] <- predict(scaleVals, X_test_study_3[, 1:39])
X_test_study_3 <- X_test_study_3 %>%
  as.matrix()
```

## Predict on test data
```{r}
pred_study_3 <- predict(fit, X = X_test_study_3,
                        type = "response")

write.csv(as.data.frame(pred_study_3),
          '../auc_data/mg_model_auc_study_3.csv')

pred_study_3 <- prediction(pred_study_3, condition_3)
pred_study_3_perf <- ROCR::performance(pred_study_3, "tpr", "fpr")


plt_dat_3 = data.frame(
  FPR = pred_study_3_perf@x.values[[1]],
  TPR = pred_study_3_perf@y.values[[1]]
)

auc_perf_study_3 <- ROCR::performance(pred_study_3, measure = "auc")
auc_study_3 <- auc_perf_study_3@y.values[[1]]
```

```{r}
# Generate AUC plot
ggplot(plt_dat_3, aes(x = FPR, y = TPR)) +
  geom_line(colour = "blue") +
  labs(x = pred_study_3_perf@x.name, y = pred_study_3_perf@y.name) +
  geom_abline(slope = 1, intercept = 0) + theme_bw() +
  theme(
    plot.title = element_text(color="Black", size=16, face="bold", hjust = 0.5),
    plot.subtitle = element_text(color = "red", size = 16, hjust = 0.5),
    axis.title.x = element_text(color="Black", size=16),
    axis.title.y = element_text(color="Black", size=16),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 16),
    strip.text.x = element_text(size = 16),
    legend.title = element_text(size=16), #change legend title font size
    legend.text = element_text(size=16) #change legend text font size)
  )  + ggtitle(paste("Overlap Group Lasso MG- Study 3 - AUC =", round(auc_study_3, 2)))
```

# Plot overlap tree
```{r}
coefs <- coef(fit) %>%
  as.matrix() %>%
  as.data.frame()
coefs$feature <- rownames(coefs)
res_mp <- MP_gLasso(fit = fit, group = fit$grp.vec, lambda.type = "min", sort.type = "max")
```


## Logit Proportions
```{r, fig.height=6, fig.width=12}
p <- plotOverlapTree(hc = overlapModel$hc_logit, coefs = coefs, type = 'logit',
                     data = res_mp$data, heatmap = F)
colnames(dat_prop) <- str_replace(colnames(dat_prop), '_prop', '')
```

```{r, fig.height=9, fig.width=10}
# generate the heatmap matrix
hmat <- generateHeatmapMatrix(df_norm, markers = useMarkers, 
                              clusters = sce_norm$mg_cell_type_distinct, threshold = 2) %>%
  as.data.frame()
rownames(hmat) <- janitor::make_clean_names(rownames(hmat))
colnames(hmat) <- janitor::make_clean_names(colnames(hmat))
```

```{r}
pdf(file = "../Plots/Manuscript_Plots/mg_tree.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 10) # The height of the plot in inches

gheatmap(p, data = hmat, offset = 0.05, hjust = 0.75, colnames_offset_y = 0, 
         color = 'white',legend_title="Proportions") +
  # vexpand(.1, -1) + 
      theme(legend.title = element_text(color = "black", size = 10),
            legend.text = element_text(color = "black", size = 10)) + 
  scale_fill_gradientn(colours = eval(formals(pheatmap::pheatmap)$color, 
                                      envir=environment(pheatmap::pheatmap)), 
                       name = "Mean Expression")
dev.off()
```

# Do some analysis
```{r, eval=FALSE}
# get non zero coefficients
colnames(coefs) <- c('Beta', 'Feature')

# get the proportion coefficients
coefs_new <- coefs[coefs$Feature %like% 'logit', ] %>%
  filter(Beta != 0)
colnames(data_logit) <- make_clean_names(colnames(data_logit))

data_logit_sub <- data_logit[, coefs_new$Feature]

sig_features <- str_replace(colnames(data_logit_sub), pattern = '_logit', '')

colnames(dat_prop) <- make_clean_names(colnames(dat_prop))
dat_prop_sub <- dat_prop[, sig_features]
```

```{r, eval=FALSE}
data_sub_final <- cbind(data_logit_sub, dat_prop_sub)
data_sub_final$Response <- condition

data_sub_final %>%
```



## Means
```{r, fig.height=6, fig.width=12}
offset <- length(overlapModel$hc_logit$data$clusters)
p.mean <- plotOverlapTree(hc = overlapModel$hc_mean, coefs = coefs, type = 'mean',
                     data = res_mp$data, heatmap = F, offset = offset)
```

```{r}
gheatmap(p.mean, data = t(hmat), offset = 0.05, hjust = 0.75, colnames_offset_y = 0, 
         color = 'white',legend_title="Proportions") +
  # vexpand(.1, -1) + 
      theme(legend.title = element_text(color = "black", size = 10),
            legend.text = element_text(color = "black", size = 10)) + 
  scale_fill_gradientn(colours = eval(formals(pheatmap::pheatmap)$color, 
                                      envir=environment(pheatmap::pheatmap)), 
                       name = "Mean Expression")
```

