---
title: "Deep learning study 4"
author: "Elijah WIllie"
date: "2024-03-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = TRUE) 
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(reticulate)
  # library(keras)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(caret)
  library(ROCR)
  library(cvAUC)
  library(spicyR)
  library(cyCombine)
  library(ClassifyR)
  library(cowplot)
  library(psych)
  library(glmnet)
  # library(ruta)
  library(ggtree)
  library(ape)
  library(grpregOverlap)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
})

# k = backend()     # this where the software used to break
source('~/Documents/PhD/bioheart_analysis/scripts/autoClass.R')
source('~/Documents/PhD/bioheart_analysis/scripts/helperFuncs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/customCV.R')
source('~/Documents/PhD/bioheart_analysis/scripts/predClusters.R')
source('~/Documents/PhD/bioheart_analysis/scripts/trainClassifier.R')
source('~/Documents/PhD/bioheart_analysis/scripts/deep_learning_helper_funcs.R')
# source('~/Documents/PhD/DeepLearning_CyTOF/robust_ae.R')
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Read in the dataset
## Study 4
```{r}
# study 4
set.seed(1994)
df <- fread('../raw_data/bioheart_ct_cytof_data_b4_mg.csv',
              nThread = 7) %>%
  as.data.frame() 

# %>%
#   group_by(sample_id) %>%
#   slice_sample(n = 5000) 

# df <- df[which(df$mg_cell_type_distinct != ""), ]
# df <- df[which(df$mg_cell_type_distinct != "other"), ]

df[, useMarkers] <- log(abs(df[, useMarkers]) + 1)

# 
preProcValues <- preProcess(df[, useMarkers], method = c("range"))
# 
df[, useMarkers] <- predict(preProcValues, df[, useMarkers])

```

# Get the reference sample for training
```{r}
res_ind <- ComputeReferenceSample(df, useMarkers)
```

# Get the encoding dimensions and the number of clusters 
```{r}
train_dat <- df[which(df$sample_id == res_ind$refSampleInd), useMarkers]
encoding_dim <- FuseSOM::computeGridSize(train_dat[, useMarkers])
```


# Define autoencoder model
```{r, echo=FALSE}
models <- autoEncoderModel(train_dat = train_dat, input_size = length(useMarkers),
                           encoding_dim = encoding_dim*2, l2_reg = 5E-04)

# model  <- robust_ae(train_dat[, useMarkers], encoding_dim = encoding_dim*2, batch_size = 16, epochs = 100)
```

# Predict on study 4
```{r}
df_norm <- predict(models$autoencoder_model, as.matrix(df[, useMarkers])) %>%
  as.data.frame()

colnames(df_norm) <- useMarkers

# df_latent <- predict(models$encoder_model, as.matrix(df[, useMarkers])) %>%
#   as.data.frame()
```

# Create SCE object for study 4 and cluster it
```{r}
sce_norm <- SingleCellExperiment(assays = list(norm = t(df_norm[, useMarkers]),
                                               raw = t(df[, useMarkers])
),
colData = df %>% dplyr::select(-useMarkers))
# sce_norm$clusters <- y_pred
```

```{r, eval=FALSE}
set.seed(1994)
sample_pts <- sample(ncol(sce_norm), 100000)
sce_sub <- sce_norm[, sample_pts]

sce_sub <- scater::runUMAP(sce_sub,
                         subset_row = useMarkers,
                         exprs_values = "raw",
                         BPPARAM = BPPARAM, 
                         n_threads = 7,
                         metric = 'cosine',
                         verbose = T)

df_raw_umap <- reducedDim(sce_sub, 'UMAP') %>%
    as.data.frame()
colnames(df_raw_umap) <- c('UMAP 1', 'UMAP 2')

df_raw_umap$Batch <- sce_sub$CyTOF.Batch

# df_raw_umap$Study <- if_else(df_raw_umap$batch %like% '3', '3', '4')
```


```{r, eval=FALSE}
sce_sub <- scater::runUMAP(sce_sub,
                         subset_row = useMarkers,
                         exprs_values = "norm",
                         BPPARAM = BPPARAM, 
                         n_threads = 7,
                         metric = 'cosine',
                         verbose = T)

df_corrected_umap <- reducedDim(sce_sub, 'UMAP') %>%
    as.data.frame()
colnames(df_corrected_umap) <- c('UMAP 1', 'UMAP 2')

df_corrected_umap$Batch <- sce_sub$CyTOF.Batch

# df_corrected_umap$Study <- if_else(df_raw_umap$Batch %like% '3', '3', '4')
```

```{r, eval=FALSE}
p.raw.umap <- df_raw_umap %>% ggplot(aes(`UMAP 1`, `UMAP 2`)) +
    geom_point(aes(color = `Batch`), size = 1) + 
    ggtitle('Raw data - Study 4') +
  theme_bw() + 
    theme(axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 15, hjust = 0.5),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12)
        ) + scale_color_d3(palette = 'category20')
```

```{r, eval=FALSE}
p.corrected.umap <- df_corrected_umap %>% ggplot(aes(`UMAP 1`, `UMAP 2`)) +
    geom_point(aes(color = `Batch`), size = 1) + 
    ggtitle('Deep learning corrected data - Study 4') +
  theme_bw() + 
    theme(axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 15, hjust = 0.5),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12)
        ) + scale_color_d3(palette = 'category20')
```

```{r, eval=FALSE}
pdf(file = "../Plots/deep_learning_UMAP_plots.pdf",   # The directory you want to save the file in
    width = 8, # The width of the plot in inches
    height = 4) # The height of the plot in inches
cowplot::plot_grid(p.raw.umap, p.corrected.umap)
dev.off()
```

## Use FuseSOM to cluster the normalised data
```{r}
# run FuseSOM clustering
nclust = 10
sce_norm <- runFuseSOM(sce_norm, numClusters = nclust, 
                       assay = 'norm',
                       verbose = FALSE)
```
## Look at concordance
```{r}
computeConcordance(sce_norm$clusters, sce_norm$mg_cell_type_distinct)
```

# Look at significant testing before and after normalisation
## Gender
```{r}
stat_test_before <- df %>%
  dplyr::select(c(useMarkers, "sample_id", "Gender", "gensini_bin")) %>%
  dplyr::group_by(sample_id, Gender, gensini_bin) %>%
  summarise_at(useMarkers, mean) %>%
  pivot_longer(-c(Gender, sample_id, gensini_bin), names_to = "variables", values_to = "value") %>%
  group_by(variables) %>%
  t_test(value ~ Gender) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()
stat_test_before
```

```{r}
df_norm$sample_id <- df$sample_id
df_norm$Gender <- df$Gender
df_norm$gensini_bin <- df$gensini_bin

stat_test_after <- df_norm %>%
  dplyr::select(c(useMarkers, "sample_id", "Gender", "gensini_bin")) %>%
  dplyr::group_by(sample_id, Gender, gensini_bin) %>%
  summarise_at(useMarkers, mean) %>%
  pivot_longer(-c(Gender, sample_id, gensini_bin), names_to = "variables", values_to = "value") %>%
  group_by(variables) %>%
  t_test(value ~ Gender) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()
stat_test_after
```

```{r}
data.frame(before_stat = stat_test_before$statistic,
                      after_stat = stat_test_after$statistic) %>%
  ggscatter(x = "before_stat", y = "after_stat",
   add = "reg.line",  # Add regressin line
   add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
   conf.int = TRUE # Add confidence interval
   ) + 
  stat_cor(aes(label = ..r.label..), label.x = 3)
```


# Compute training features
## Proportions
```{r, eval=FALSE}
# do prediction
data <- getProp(sce_norm,
                feature = "clusters", imageID = "sample_id", logit = F)

dat_prop <- data
data <- 2*asin(sqrt(data))

clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

row_names <- rownames(data)

Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"]
GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")]

clinicaldata_info <- cbind(Age, GenderM)

condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])


data_logit <- getProp(sce_norm,
                feature = "clusters", imageID = "sample_id", logit = T)
colnames(data_logit) <- paste0(colnames(data_logit), "_logit")
colnames(data) <- paste0(colnames(data), "_prop")
colnames(dat_prop) <- colnames(data)
```

## Mean expression
```{r, eval=FALSE}
t <- df_norm
t$sample_id <- df$sample_id

data_mean <- t %>% 
  dplyr::group_by(sample_id) %>%
  dplyr::summarise_at(vars(useMarkers), mean) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

data_median <- t %>% 
  group_by(sample_id) %>%
  summarise_at(vars(useMarkers), median) %>%
  dplyr::select(-sample_id) %>%
  as.data.frame()

rownames(data_mean) <- clinicaldata$sample_id
rownames(data_median) <- clinicaldata$sample_id
colnames(data_mean) <- paste0(colnames(data_mean), "_mean")
colnames(data_median) <- paste0(colnames(data_median), "_median")
```

# Look at CV performance
```{r, eval=FALSE}
res.cv <- KFoldCustomEnsemble(dat_prop = data, dat_logit = data_logit, 
                              dat_mean = data_mean,
                              train_y = condition, alpha = 0.5, 
                              nRepeats = 10, k = 10)
```

```{r, eval=FALSE}
aucs <- data.frame(`DeepLasso` = res.cv$cv_res)
write.csv(aucs, file = '../auc_data/deep_learning_study_4_auc.csv', row.names = F)
```

```{r, fig.height=6, fig.width=9, eval=FALSE}
# pdf(file = "Plots/SVM_CV_plot.pdf",   # The directory you want to save the file in
#     width = 8, # The width of the plot in inches
#     height = 6) # The height of the plot in inches

cv_plot_dat <- res.cv$cv_dat
out <- cvAUC(cv_plot_dat$predictions, cv_plot_dat$labels)

plot(out$perf, col = "black", lty = 3, size = 2,
     main = paste0("10-fold CV with 10 repeats. Average AUC = ",
                                                      round(out$cvAUC, 2)))
plot(out$perf, col ="red", avg = "threshold", add = TRUE) 

# dev.off()
```

# Look at batch correction performance
```{r, eval=FALSE}
labels <- df_norm %>%
          cyCombine::create_som(rlen = 10,
                                xdim = 8,
                                ydim = 8,
                                markers = useMarkers)

# Add labels
df_norm <- df_norm %>%
  dplyr::mutate(som = labels)
df_norm$id <- 1:nrow(df_norm)
df_norm$batch <- df$CyTOF.Batch


# Transfer labels to uncorrected data
df$id <- 1:nrow(df_norm)
df <- df_norm %>%
  dplyr::select('id', all_of('som')) %>%
  dplyr::left_join(df, by = "id")

df$batch <- df$CyTOF.Batch
```

## Look at Median Absolute Deviation (MAD) score
```{r, eval=FALSE}
mad_val <- df %>%
      cyCombine::evaluate_mad(df_norm,
                              filter_limit = NULL,
                              markers = useMarkers,
                              cell_col = 'som')

mad_dat <- mad_val$mad
```

## Generate histogram plot for batch correction
```{r, eval=FALSE}
pdf(file = "../Plots/deep_learning_MAD_plot.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 4) # The height of the plot in inches
mad_dat %>%
  ggplot(aes(x = Difference)) + 
  geom_histogram(bins = 30) +
  ggtitle('Histogram of MAD scores for Deep Learning', 
          subtitle = paste('Median score:', mad_val$score)) +
  theme_bw() + 
    theme(axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 15, hjust = 0.5,
                                  face = 'bold'),
        plot.subtitle = element_text(size = 15, 
                                     hjust = 0.5, colour = 'blue'),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12)
        )
dev.off()
```

