---
title: "scMerge2 all batches"
author: "Elijah WIllie"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = TRUE) 
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(caret)
  library(ROCR)
  library(cvAUC)
  library(cowplot)
  library(scMerge)
})
source('~/Documents/PhD/bioheart_analysis/scripts/helperFuncs.R')
source('~/Documents/PhD/bioheart_analysis/scripts/deep_learning_helper_funcs.R')

# source('~/Documents/PhD/DeepLearning_CyTOF/robust_ae.R')
```

```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")

final_celltypes <- c(
   "B cells",
   "CD8hi",
   "14+ monos",
   "CD4+ Tconv",
   "CD8lo",
   "NK",
   "16+ monos",
   "CD4+ Treg",
   "pDCs",
   "CD141+ DCs"
    )
```

# Read in the datasets
```{r}
set.seed(1994)
df_4 <- fread('../raw_data/bioheart_ct_cytof_data_b4_mg.csv',
              nThread = 7) %>%
  as.data.frame() %>%
  dplyr::select(c(useMarkers, "sample_id", "CyTOF.Batch", "mg_cell_type_distinct"))

# df_4 <- df_4[which(df_4$mg_cell_type_distinct != ""), ]

colnames(df_4)[colnames(df_4) == "CyTOF.Batch"] = "batch"
colnames(df_4)[colnames(df_4) == "mg_cell_type_distinct"] = "CellType"

```

```{r}
df_3 <- fread('../Study_3_2019/all_batches_processed_mg_10K.csv',
              nThread = 7) %>%
  as.data.frame() %>%
  dplyr::select(c(useMarkers,  "sample_id", "CellTypes", "Batch"))

colnames(df_3)[colnames(df_3) == "Batch"] = "batch"
colnames(df_3)[colnames(df_3) == "CellTypes"] = "CellType"
```

## Combine study 4 and study 3
```{r}
df <- rbind(df_4, df_3) %>%
  dplyr::filter(CellType %in% final_celltypes)

# update study information
# df$batch <- if_else(df$batch_old %like% 4, 4, 3)
```


```{r}
rowsum_df <- DelayedMatrixStats::rowSums2(df[, useMarkers] %>% as.matrix())
df <- df[rowsum_df != 0, ]
```

```{r}
df[, useMarkers] <- cyCombine::transform_asinh(df, markers = useMarkers)
```

# Batch correct using scMerge2
```{r}
sce <- SingleCellExperiment(assays = list(counts = t(df[, useMarkers])
),
colData = df %>% dplyr::select(-useMarkers))
```


```{r}
exprsMat <- counts(sce)
colnames(exprsMat) <- rownames(df)
scMerge2_res <- scMerge2(exprsMat = exprsMat,
                         batch = sce$batch,
                         ctl = useMarkers,
                         verbose = TRUE,
                         ruvK = 5,
                         seed = 1994)
```

```{r}
colnames(scMerge2_res$newY) <- NULL
assay(sce, "scMerge2", withDimnames=FALSE) <- scMerge2_res$newY
df_norm <- scMerge2_res$newY %>%
  t() %>%
  as.data.frame()
```

# Cluster into 12 clusters
```{r}
# run FuseSOM clustering
nclust = 10
sce$clusters_scMerge2 <- NULL
sce <- runFuseSOM(sce, numClusters = nclust, assay = 'scMerge2',
                       verbose = FALSE, clusterCol = "clusters_scMerge2")
```

# Compute concordance on full data
```{r}
computeConcordance(sce$clusters_scMerge2, sce$CellType)
```

# Train cell type model on all batches
## Select training data
```{r}
set.seed(1994)
df_norm$CellType <- factor(sce$CellType)
df_norm$sample_id <- sce$sample_id
df_norm$batch <- sce$batch
```

# Test on batch 3
## MG
```{r}
train_batch <- df_norm[df_norm$batch %like% '4', ] %>%
  dplyr::mutate(cellTypes = CellType) %>%
  dplyr::select(c(useMarkers, cellTypes))
test_batch <- df_norm[df_norm$batch %like% '3', ] %>%
  dplyr::select(c(useMarkers, CellType))
```

## Select nop N sample
```{r}
res_ind <- ComputeReferenceSample(train_batch, markers = useMarkers, N = 2)
train_batch_sub <- train_batch[which(train_batch$sample_id %in% res_ind$topNSamples), ] %>%
  dplyr::select(-sample_id)
```


```{r}
test_celltypes <- cellTypeClassifier(train_x = train_batch, 
                                     test_x = test_batch[, useMarkers], model = 'lda')
```

```{r}
computeConcordance(test_batch$CellType, test_celltypes)
```

```{r}
MLmetrics::Accuracy(test_celltypes, test_batch$CellType)
```

```{r}
MLmetrics::F1_Score(test_celltypes, test_batch$CellType)
```

```{r}
classifier_metrics <- ml_test(test_celltypes, test_batch$CellType, output.as.table = F)
classifier_metrics$balanced.accuracy
classifier_metrics$F1
```

```{r}
classifier_metrics_tab <- ml_test(test_celltypes, test_batch$CellType, output.as.table = T) %>%
  dplyr::select(balanced.accuracy, F1) %>%
  rownames_to_column(var = "Celltype")

```


# Generate classification bar-chart
```{r}
conf_matrix <- tibble("Truth" = test_batch$CellType,
                    "Predicted" = test_celltypes) %>%
  table()

data <- as.data.frame(melt(conf_matrix))
names(data) <- c("Truth", "Predicted", "Count")

# Calculate the total support for each actual class
total_support <- aggregate(Count ~ Truth, data = data, sum)

# Merge total support back to the original data for proportions
data <- merge(data, total_support, by = "Truth", all.x = TRUE)

names(data)[names(data) == "Count.x"] <- "Count"
names(data)[names(data) == "Count.y"] <- "TotalSupport"

# Calculate proportion
data$Proportion <- data$Count / data$TotalSupport

# Creating the plot
ggplot(data, aes(x = Truth, y = Proportion, fill = Predicted)) +
  geom_bar(stat = "identity") +
  labs(title = "scMerge2 - Class Support and Prediction Proportions",
       x = "Truth",
       y = "Proportion of Total Support",
       fill = "Predicted") +
  theme_minimal() + 
  theme_bw() +
    theme(axis.text.x = element_text(size = 12),
        plot.title = element_text(size = 15, hjust = 0.5),
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12)
        ) + scale_color_d3(palette = 'category20') + 
  scale_fill_aaas()
```

# write predicted data to file
```{r}
# write balanced accuracy and F1
write.csv(classifier_metrics_tab, '../concordance_data/scMerge2_batch_3_pred_acc.csv',
          row.names = F)

# write plot data
write.csv(data, '../concordance_data/scMerge2_batch_3_pred_data.csv',
          row.names = F)
```

# Test on each batch 
## MG
```{r, eval=FALSE}
batches <- unique(df_norm$batch)
ari_mg <- c()
nmi_mg <- c()
for(batch in batches){
  set.seed(1994)
  message(paste0("Now prediction on batch: ", batch))
  df_batch <- df_norm[df_norm$batch == batch, ]
  df_norm_sub <- df_norm[df_norm$batch != batch, ] %>%
  group_by(clusters_mg) %>%
  slice_sample(n = 1000)
  nnet_model_mg <- fitCelltypeModel(train_x = df_norm_sub[, useMarkers],
                                  train_y = df_norm_sub$clusters_mg)
  batch_clusters_mg <- nnet_model_mg %>%
    predict(as.matrix(df_batch[, useMarkers]))
  pred_clusters_mg <- batch_clusters_mg$predicted
  concor_mg <- computeConcordance(df_batch$clusters_mg, pred_clusters_mg)
  ari_mg <- c(ari_mg, concor_mg$ARI)
  nmi_mg <- c(nmi_mg, concor_mg$NMI)
}
```

## FuseSOM clusters
```{r, eval=FALSE}
ari_fs <- c()
nmi_fs <- c()
for(batch in batches){
  set.seed(1994)
  message(paste0("Now prediction on batch: ", batch))
  df_batch <- df_norm[df_norm$batch == batch, ]
  df_norm_sub <- df_norm[df_norm$batch != batch, ] %>%
  group_by(clusters_fs) %>%
  slice_sample(n = 1000)
  nnet_model_fs <- fitCelltypeModel(train_x = df_norm_sub[, useMarkers],
                                  train_y = df_norm_sub$clusters_fs)
  batch_clusters_fs <- nnet_model_fs %>%
    predict(as.matrix(df_batch[, useMarkers]))
  pred_clusters_fs <- batch_clusters_fs$predicted
  concor_fs <- computeConcordance(df_batch$clusters_fs, pred_clusters_fs)
  ari_fs <- c(ari_fs, concor_fs$ARI)
  nmi_fs <- c(nmi_fs, concor_fs$NMI)
}
```

# Collate the final data
```{r, eval=FALSE}
dat_final_mg <- data.frame(Batch = batches, ARI = ari_mg, 
                           NMI = nmi_mg, Cluster = rep('MG', length(batches)))
dat_final_fs <- data.frame(Batch = batches, ARI = ari_fs, 
                           NMI = nmi_fs, Cluster = rep('FuseSOM', length(batches)))

dat_final <- rbind(dat_final_mg, dat_final_fs)

write.csv(dat_final, '../concordance_data/scMerge2_all_batches.csv', row.names = F)
```

# Save SCE object
```{r, eval=T}
save(sce, file = "../sce_dat/bioheart_scMerge2_batch_corrected.RData")
```


