---
title: "CMV Prediction"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tidyverse)
  library(FuseSOM)
  library(data.table)
  library(SingleCellExperiment)
  library(dioscRi)
})
```

```{r}
split_and_extract <- function(string, n) {
  split_strings <- strsplit(string, "\\.")[[1]][[2]]  # Split strings by comma
  return(split_strings)
}
```


```{r}
metaData <- read.csv('~/Documents/PhD/DeepLearning_CyTOF/DeepLearningCyTOF/cmv_metadata.csv')
metaData <- metaData[, -1]

metaData$sample_id <- lapply(metaData$name, split_and_extract)
metaData$sample_id <- unlist(metaData$sample_id)

useMarkers <- c("TCRGD", "IGD","HLADR", "CD94","CD85J","CD8","CD56",
                "CD45RA","CD4","CD38", "CD33", "CD3","CD28","CD27",
                "CD25","CD24","CD20","CD19","CD161","CD16","CD14","CD127","CCR7")
```

```{r}
load("~/Documents/PhD/DeepLearning_CyTOF/DeepLearningCyTOF/Data/cmv_train_MMD_VAE_sce.RData")
load("~/Documents/PhD/DeepLearning_CyTOF/DeepLearningCyTOF/Data/cmv_test_MMD_VAE_sce.RData")
```


## cluster training data
```{r, eval=F}
# run FuseSOM clustering
sce_train$clusters <- NULL
sce_test$clusters <- NULL
nclust = 11
sce_train <- runFuseSOM(sce_train, numClusters = nclust, assay = 'norm',
                       verbose = F)
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_train, type = "VAE") %>%
  mutate(cellTypes = sce_train$clusters)

test_x <- reducedDim(sce_test, type = "VAE")
```

```{r, eval=T}
test_clusters <- trainCellTypeClassifier(trainX = train_x, 
                                         testX = test_x, model = 'lda')
table(test_clusters)
sce_test$clusters <- test_clusters
```

# Set up clinical data
## Training
```{r}
clinicaldata <- colData(sce_train) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, CMV_Ab) %>%
  distinct()
clinicaldata$condition <- factor(if_else(clinicaldata$CMV_Ab == "True", 1, 0))
```

## Testing
```{r}
clinicaldata_test <- colData(sce_test) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, CMV_Ab) %>%
  distinct()
clinicaldata_test$condition <- factor(if_else(clinicaldata_test$CMV_Ab == "True", 1, 0))
```


# Compute training features
## Proportions
```{r}
data_logit <- computeFeatures(sce = sce_train, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "condition"])
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_train, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

## Compute features on test set
## Proportions
```{r}
data_test_logit <- computeFeatures(sce = sce_test, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_test <- rownames(data_test_logit)
condition_test <- factor(clinicaldata_test[match(row_names_test,clinicaldata_test$sample_id),
                                      "condition"])
```

## Means
```{r, eval=T}
markerMeanCellType_test <- computeFeatures(sce = sce_test, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

# Generate grouping structures
```{r}
nclust <- length(unique(test_clusters))
tree <- generateTree(features = data_logit, method = "ward")
tree <- tree$tree
groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)

order <- tree$order
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
  
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_test_logit, markerMeanCellType_test) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```


## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
y_test <- as.numeric(levels(condition_test))[condition_test]
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Test set AUC =")
test_auc$plot

# write.csv(test_auc$preds,
#           'cmv_dioscRi_aucs.csv')
```


# Visualise Resulting Tree
```{r, fig.height=10, fig.width=18}
visualiseModelTree(fit = fit$fit, tree = tree, type = "cluster", 
                   trainingData = X_train)
```