---
title: "Bioheart Prediction"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
library(dioscRi)
set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```
# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
cluster_col <- "clusters_norm"
```


# Load normalised data
```{r}
load('../sce_dat/study_4_MMD_VAE_updated.RData')
load('../sce_dat/study_3_MMD_VAE_updated.RData')
```

## Use FuseSOM to cluster the normalised data
```{r, eval=T}
# run FuseSOM clustering
nclust = 11
sce_norm$cluster_col <- NULL
sce_norm <- runFuseSOM(sce_norm, numClusters = nclust, assay = 'norm',
                       verbose = FALSE, clusterCol = 'clusters_norm')
# sce_norm$clusters <- sce_temp$clusters
```

# Annotate clustering with Helen's annotations
```{r, eval=T}
# sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
# sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != ""]

sce_norm$clusters_norm <- recode_factor(sce_norm$clusters_norm,
                                 "cluster_1" = "NK", # NK Cells
                                 "cluster_2" = "Not Naive CD8+ T cells", # Not Naive CD8+ T cells
                                 "cluster_3" = "Naive CD8+ T cells", # Naive CD8+ T cells
                                 "cluster_4" = "NK/Mixed", # NK Cells
                                 "cluster_5" = "Naive B Cells", # mix of monocytes and dendritic cells
                                 "cluster_6" = "CD4+ Tconv", # mix of naïve CD4+ T cells, B cells and NK cells
                                 "cluster_7" = "Memory B cells", # B cells
                                 "cluster_8" = "16+ monocytes", # NK cells
                                 "cluster_9" = "CD141+ DC/pDCs", # central memory CD8+ T cells
                                 "cluster_10" = "14+ monocytes", # naïve CD8+ T cells
                                 "cluster_11" = "CD4+ Treg" # terminal effector memory revertant CD8+ T cells
                                 )
```

# Look at the clustering performance per sample
```{r}
df_norm <- assay(sce_norm, type = "norm") %>%
  t() %>%
  as.data.frame() %>%
  # mutate(cellTypes = sce_norm$mg_cell_type_distinct) %>%
  mutate(predicted_norm = sce_norm$clusters_norm)
  # mutate(predicted_raw = sce_norm$clusters_raw) %>%
  # mutate(sample_id = sce_norm$sample_id)
```

```{r}
plotExpBox <- function(dat, cluster = NULL){
  p <- dat %>%
  dplyr::filter(predicted_norm == cluster) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(outliers = F) +
    geom_jitter(alpha = 0.1) +
  xlab("Marker") + 
  ylab("Expression") +
    scale_y_continuous(limits = c(0,1)) +
  facet_wrap(~predicted_norm, scales = "free") +
     theme(legend.title = element_text(color = "black", size = 16),
          plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
          legend.text = element_text(color = "black", size = 16),
          axis.text.x = element_text(angle = 90, vjust = 0.5, 
                                     hjust=1, size = 16, face = "bold"),
          axis.text.y = element_text(size = 16, face = "bold"),
          strip.text.x = element_text(size = 20, colour = "orange"),
          axis.title.x = element_text(color="Black", size=16),
      axis.title.y = element_text(color="Black", size=16))
  
  plot(p)
}
```


```{r, eval=FALSE}
pdf(file = "../Plots/All_expression_plot.pdf",   # The directory you want to save the file in
    width = 16, # The width of the plot in inches
    height = 10) # The height of the plot in inches
df_norm_melted %>%
  # dplyr::filter(predicted_norm == cluster) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(outliers = F) +
    geom_jitter(alpha = 0.1) +
  xlab("Marker") + 
  ylab("Expression") +
  ggtitle("All expression") +
    scale_y_continuous(limits = c(0,1)) +
  # facet_wrap(~predicted_norm, scales = "free") +
     theme(legend.title = element_text(color = "black", size = 16),
          plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
          legend.text = element_text(color = "black", size = 16),
          axis.text.x = element_text(angle = 90, vjust = 0.5, 
                                     hjust=1, size = 16, face = "bold"),
          axis.text.y = element_text(size = 16, face = "bold"),
          strip.text.x = element_text(size = 20, colour = "orange"),
          axis.title.x = element_text(color="Black", size=16),
      axis.title.y = element_text(color="Black", size=16))

dev.off()
```


```{r, eval=FALSE}
set.seed(1994)
df_norm_melted <- melt(df_norm) %>%
  dplyr::group_by(predicted_norm, variable) %>%
  slice_sample(n = 1000)

clusters <- unique(df_norm_melted$predicted_norm) %>%
  as.character()


p <- ggplotify::as.ggplot(scater::plotGroupedHeatmap(sce_norm, 
                           features = useMarkers, 
                           group = cluster_col, 
                           block = cluster_col,
                           exprs_values = "norm",
                           center = TRUE, 
                           scale = TRUE, 
                           zlim = c(-2,2),
                           cluster_rows = FALSE)) + 
  # theme_bw() +
  # theme_minimal() +
  ggtitle(label = "Heatmap of normalised clusters expression") +
    theme(
          plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
    )

pdf(file = "../Plots/Clusters_13_expression_plots.pdf",   # The directory you want to save the file in
    width = 16, # The width of the plot in inches
    height = 10) # The height of the plot in inches

plot(p)

for(cluster in clusters){
  plotExpBox(df_norm_melted, cluster = cluster)
}

dev.off()
```


```{r, eval=FALSE}
num_samples <- length(unique(df_norm$sample_id))
samples <- unique(df_norm$sample_id)
ari_norm <- c()
ari_raw <- c()
 for (i in 1:num_samples) {
    cellTypes <- df_norm[df_norm['sample_id'] == samples[[i]], 'cellTypes']
    predicted_norm <- df_norm[df_norm['sample_id'] == samples[[i]], 'predicted_norm']
    predicted_raw <- df_norm[df_norm['sample_id'] == samples[[i]], 'predicted_raw']
    
    ARI_norm <- computeConcordance(cellTypes, predicted_norm)$ARI %>% round(2)
    ARI_raw <- computeConcordance(cellTypes, predicted_raw)$ARI %>% round(2)
    ari_norm <- c(ari_norm, ARI_norm)
    ari_raw <- c(ari_raw, ARI_raw)
 }

df_ari <- data.frame(sample_id = as.factor(samples), norm = ari_norm, raw = ari_raw,
           diff = ari_norm - ari_raw)
```


```{r, fig.height=10, fig.width=16, eval=FALSE}
ggplotify::as.ggplot(scater::plotGroupedHeatmap(sce_norm, 
                           features = useMarkers, 
                           group = cluster_col, 
                           block = cluster_col,
                           exprs_values = "norm",
                           center = TRUE, 
                           scale = TRUE, 
                           zlim = c(-2,2),
                           cluster_rows = FALSE)) + 
  # theme_bw() +
  # theme_minimal() +
  ggtitle(label = "Heatmap of normalised clusters expression") +
    theme(
          plot.title = element_text(size = 20, hjust = 0.5, face = "bold")
    )
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_norm, type = "VAE") %>%
  dplyr::mutate(cellTypes = as.factor(sce_norm$clusters_norm))

test_x <- reducedDim(sce_3_norm, type = "VAE")
```

```{r, eval=T}
df_3_clusters <- trainCellTypeClassifier(trainX = train_x, testX = test_x, model = 'lda')

table(df_3_clusters)

sce_3_norm$clusters <- as.factor(df_3_clusters)
```

# Save SCE objects
```{r, eval=F}
save(sce_norm, file = '../../sce_dat/study_4_MMD_VAE_updated.RData')
save(sce_3_norm, file = '../../sce_dat/study_3_MMD_VAE_updated.RData')
```

# Compute training features
## Proportions
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)
# 
# row_names <- clinicaldata$sample_id
# 
# Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       "Age"]
# GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       c("GenderM")]
# 
# clinicaldata_info <- cbind(Age, GenderM)
```

## Generate features
### Proportions
```{r}
clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
# clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)
clinicaldata_3$Gensini_bin <- as.factor(clinicaldata_3$Gensini_bin)
```


# Compute training features
## Proportions
```{r}
data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = F, useMarkers = useMarkers)

row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])
# colnames(data_logit) <- as.factor(colnames(data_logit))

Age <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"])
GenderM <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")])

clinicaldata_info <- cbind(Age, GenderM)

# data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
#                               cellTypeCol = 'cluster_col', sampleCol = 'sample_id', 
#                               logit = F, useMarkers = useMarkers)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))

# markerMeanCellType_rsc <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = 'norm_rescaled',
#                               cellTypeCol = 'cluster_col', sampleCol = 'sample_id', 
#                               logit = T, useMarkers = useMarkers)
# colnames(markerMeanCellType_rsc) <- as.factor(colnames(markerMeanCellType_rsc))
```


## Compute features on study 3
## Proportions
```{r}
data_3_logit <- computeFeatures(sce = sce_3_norm, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_3 <- rownames(data_3_logit)
condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])
colnames(data_3_logit) <- as.factor(colnames(data_3_logit))

Age <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType_3 <- computeFeatures(sce = sce_3_norm, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))
```


# Generate grouping structures
```{r}
nclust <- length(unique(sce_norm$clusters_norm))
trees <- generateTree(features = data_logit, method = "ward")
tree <- trees$tree
order <- trees$order
groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Add the clinical data in
```{r, eval=F}
groups[[length(groups) + 1]] <- 309:310
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
# X_train <- cbind(X_train, clinicaldata_info)
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_3_logit, markerMeanCellType_3) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
# X_test <- cbind(X_test, clinicaldata_info_3)
y_test <- as.numeric(levels(condition_3))[condition_3]
```


## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, modelSummary = F,
                groups = groups, penalty = "grLasso", seed = 1994)
```

```{r}
train_auc <- plotAUC(fit = fit$fit, xTest = X_train, yTest = y_train, title = "Study 3 - All AUC =")
train_auc$plot
```


## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = condition_3, title = "Study 3 - All AUC =")
test_auc$plot

# write.csv(test_auc$preds,
#           '../auc_data/fs_dioscRi_auc_all_study_3.csv')

```

# Visualise Resulting Tree
```{r, fig.height=20, fig.width=40}
p.all <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG",
                   trainingData = X_train, title = "Unsupervised Clustering")

pdf(file = "../dioscRi_manuscript/Plots/Supp_Figure_2_tree_full_unsupervised.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 12) # The height of the plot in inches
    p.all 
dev.off()
# ggsave(file="../Plots/fs_tree.svg", plot=p, width=18, height=16, dpi = 1200)
```

```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_mean$sigFeatures
stats <- stats_mean$stats



plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin")

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```

```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop <- getSigFeatures(fit$fit, type = 'prop',  prop = data,
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_prop$sigFeatures
stats <- stats_prop$stats



plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin")

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```

```{r,fig.width=28, fig.height=16, eval=F}
pdf(file = "../Plots/Manuscript_Plots/Figure_4.pdf",   # The directory you want to save the file in
    width = 36, # The width of the plot in inches
    height = 18) # The height of the plot in inches
cowplot::plot_grid(p.tree.mg, p.tree, labels = LETTERS[1:2])

dev.off()
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit %>%
  as.matrix()
```

## Testing
```{r}
X_test <- data_3_logit %>%
  as.matrix()

```

```{r}
groups <- generateGroups(tree = tree, nClust = 11, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train,
                groups = groups, penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot

# write.csv(test_auc$preds,
#           '../auc_data/fs_dioscRi_auc_prop_study_3.csv')

```
## Look at tree plot
```{r, fig.height=20, fig.width=40}
p.prop <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", 
                   trainingData = X_train, heatmap = FALSE)

pdf(file = "../dioscRi_manuscript/Plots/Supp_Figure_2_tree_prop_unsupervised.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 12) # The height of the plot in inches
    p.prop 
dev.off()
```

```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_prop <- getSigFeatures(fit$fit, type = 'prop',  prop = data,
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_prop$sigFeatures
stats <- stats_prop$stats



plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin")

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType %>%
  as.matrix()
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_3 %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train,
                groups = groups, penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot

# write.csv(test_auc$preds,
#           '../auc_data/fs_dioscRi_auc_means_study_3.csv')

```


```{r, fig.height=24, fig.width=40}
p.means <- plotHeatmap(fit$fit, order = order, type = "MG", markerOrder = )
```

```{r, fig.width=20, fig.height=12, eval=T}
pdf(file = "../dioscRi_manuscript/Plots/Supp_Figure_2_tree_means_unsupervised.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches
    p.means 
dev.off()
```

```{r, eval=T, fig.height=16, fig.width=20}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_mean$sigFeatures
# stats <- stats_mean$stats[stats_mean$stats$p < 0.05, ]



plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin")

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```

## Study 3
```{r, eval=T, fig.height=10, fig.width=16, eval=FALSE}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean_3 <- getSigFeatures(fit, type = 'mean',  
                         mean = markerMeanCellType_3, clinicaldata = clinicaldata_3, 
                         outcome = "Gensini_bin")

if(sum(stats_mean_3$stats$p.adj < 0.05) > 0){
  plotSigFeatures(stats = stats_mean_3$stats, sigFeatures = stats_mean_3$sig_features, outcome = "Gensini bin")
}
```

# look at boxplots
```{r}
# Step 1: Add combined value + bin label
data <- data %>%
  dplyr::mutate(`NK_NK/mixed` = rowSums(across(c("NK_logit", "NK/Mixed_logit"))),
         Gensini_bin = as.factor(y_train))

# Step 2: Reshape to long format and clean facet names
data_long <- data %>%
  pivot_longer(
    cols = c("NK_logit", "NK/Mixed_logit", "NK_NK/mixed"),
    names_to = "Celltype",
    values_to = "Logit_value"
  ) %>%
  mutate(Celltype = str_remove(Celltype, "_logit"))  # Remove suffix

# Step 3: Plot
ggplot(data_long, aes(x = Gensini_bin, y = Logit_value)) +
  geom_boxplot() +
  facet_wrap(~ Celltype, scales = "free_y") +
  xlab("Gensini") +
  ylab("Proportion") +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(hjust = 1),
    axis.title = element_text(size = 12)
  )
```
```{r}
ggplot(data, aes(x = NK_logit, y = `NK/Mixed_logit`, colour = Gensini_bin)) +
  geom_point() +
  # facet_wrap(~ Celltype, scales = "free_y") +
  xlab("NK_logit") +
  ylab("NK/Mixed_logit") +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(hjust = 1),
    axis.title = element_text(size = 12)
  )
```

