---
title: "Bioheart Prediction Manual Gating"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
library(dioscRi)
library(ggplot2)
library(FuseSOM)
library(SingleCellExperiment)
library(tidyverse)
library(data.table)
library(data.tree)
tensorflow::set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA-DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c-PE')
```




# Load normalised data
```{r}
bioheart_root <- Sys.getenv(
  "BIOHEART_ROOT",
  unset = "~/Documents/Academic/PhD/bioheart_analysis"
)
sce_dir <- file.path(bioheart_root, "data", "dioscRi_analysis_data",
                     "sce_objects", "BioHEART")
load(file.path(sce_dir, "study_4_MMD_VAE_updated.RData"))
load(file.path(sce_dir, "study_3_MMD_VAE_updated.RData"))
```

```{r}
df_raw <- cbind(colData(sce_norm), t(as.matrix(assay(sce_norm, "raw")))) %>%
    as.data.frame()
```


# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_norm, type = "VAE") %>%
  dplyr::mutate(cellTypes = sce_norm$mg_cell_type_distinct)


test_x <- reducedDim(sce_3_norm, type = "VAE") %>%
  as.data.frame()
```

## Fit SVM model
```{r, eval=T}
df_3_clusters <- dioscRi::trainCellTypeClassifier(trainX = train_x, 
                                                  testX = test_x, model = 'lda')

table(df_3_clusters)

sce_3_norm$mg_clusters <- df_3_clusters
```

# Setup Clinical information
## Study 4
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

```

## Study 3
```{r}
clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)
clinicaldata_3$Gensini_bin <- as.factor(clinicaldata_3$Gensini_bin)
```


# Compute training features
## Proportions
```{r}
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != ""]

data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', 
                              sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])

data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = F, useMarkers = useMarkers)
Age <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"])
GenderM <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")])

clinicaldata_info <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = "norm",
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```


## Compute features on study 3

## Proportions
```{r}
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != "other"]
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != ""]
sce_3_norm$mg_clusters <- droplevels(sce_3_norm$mg_clusters)

data_3_logit <- computeFeatures(sce = sce_3_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_3 <- rownames(data_3_logit)
condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])

Age <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType_3 <- computeFeatures(sce = sce_3_norm, featureType = 'mean',  assay = "norm",
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

# Generate Tree
```{r, eval=T}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_logit')
    `14+ monos` <- myeloid$AddChild('14+ monos_logit')
    `16+ monos` <- myeloid$AddChild('16+ monos_logit') 
    pDCs <- myeloid$AddChild('pDCs_logit')
    `CD141+ DCs` <- myeloid$AddChild('CD141+ DCs_logit')
    `CD1c+ DCs` <- myeloid$AddChild('CD1c+ DCs_logit')
  NK <- root$AddChild('NK_logit')
  `B cells` <- root$AddChild('B cells_logit')
  CD3 <- root$AddChild('CD3_logit')
    CD8hi <- CD3$AddChild('CD8hi_logit')
    CD8lo <- CD3$AddChild('CD8lo_logit')
    CD4 <- CD3$AddChild('CD4_logit')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_logit')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_logit')
  # other <- root$AddChild('other_logit')
```



# Generate grouping structures
```{r}
nclust <- ncol(data_logit)
order <- ape::as.phylo(root)$tip.label
tree <- ggtree::ggtree(ape::as.phylo(root),ladderize = F, layout = 'dendrogram')

tree <- dioscRi::findChildren(tree)

groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Add the clinical data in
```{r, eval=F}
groups[[length(groups) + 1]] <- 309:310
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
 # Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
# X_train <- cbind(X_train, clinicaldata_info)
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_3_logit, markerMeanCellType_3) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
# X_test <- cbind(X_test, clinicaldata_info_3)
y_test <- as.numeric(levels(condition_3))[condition_3]
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994, alpha = NULL)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot
```



# Visualise Resulting Tree
```{r, fig.height=10, fig.width=12}
visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", trainingData = X_train,
                   nodesToRemove =  c("Root", "Myeloid", "CD3", "CD4"), title = "Proportions + Marker Means")
```

```{r, eval=T, fig.height=10, fig.width=16}
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_mean$sigFeatures
stats <- stats_mean$stats

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```


# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test<- data_3_logit %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- generateGroups(tree = tree, nClust = nclust, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot
```

## Look at tree plot
```{r, fig.height=5, fig.width=6}
visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", trainingData = X_train,
                   nodesToRemove =  c("Root", "Myeloid", "CD3", "CD4"), heatmap = FALSE)
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_3 %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot
```


```{r, fig.height=5, fig.width=6}
order <- gsub("_logit", "", order)
order <- gsub("_", " ", order)
plotHeatmap(fit$fit, type = "MG", order = order)

```