---
title: "BioHEART Prediction iMUBAC"
author: "Elijah Willie"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(keras3)
  library(tensorflow)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(caret)
  library(ROCR)
  library(psych)
  library(tensorflow)
  library(glmnet)
  library(data.tree)
  library(ggtree)
  library(ape)
  library(grpreg)
  library(treekoR)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
})
files.sources = list.files(path = '..//VaeTOF/VaeTOF/R/', full.names = T)
sapply(files.sources, source)
source('~/Documents/PhD/bioheart_analysis/old_scripts/estimate.ED.R')
set_random_seed(1994)
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```
# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
```

# Load the imubac normalised data
```{r}
load('../sce_dat/bioheart_iMUBAC_batch_corrected.RData')
```

## Subset the data
```{r, eval=T}
sce_imubac_disc <- sce_imubac[, sce_imubac$batch == "4"]


sce_imubac_val <- sce_imubac[, sce_imubac$batch == "3"]

```

## Use FuseSOM to cluster the normalised data
```{r, eval=F}
# run FuseSOM clustering
nclust = 11
sce_imubac_disc$clusters_norm <- NULL
sce_imubac_disc <- runFuseSOM(sce_imubac_disc, numClusters = nclust, 
                                 assay = 'normexprs',
                       verbose = FALSE, clusterCol = 'clusters_norm')
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- assay(sce_imubac_disc, "normexprs") %>%
    t() %>%
    as.data.frame() %>%
  mutate(cellTypes = as.factor(sce_imubac_disc$CellType))

test_x <- assay(sce_imubac_val, "normexprs") %>%
    t() %>%
    as.data.frame()
```

```{r, eval=T}
val_clusters <- cellTypeClassifier(trainData = train_x, testData = test_x, model = 'lda')

table(val_clusters)

sce_imubac_val$mg_clusters <- as.factor(val_clusters)
```

# Compute training features
## Proportions
```{r}
clinicaldata <- colData(sce_imubac_disc) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

```

## Generate features
### Proportions
```{r}
clinicaldata_val <- colData(sce_imubac_val) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()

encoded_gender_val <- model.matrix(~Gender-1, data=clinicaldata_val)
clinicaldata_val <- cbind(clinicaldata_val, encoded_gender_val)
clinicaldata_val$Gensini_bin <- as.factor(clinicaldata_val$Gensini_bin)
```

# Compute training features
## Proportions
```{r}
sce_imubac_disc <- sce_imubac_disc[, sce_imubac_disc$CellType != "other"]
sce_imubac_disc <- sce_imubac_disc[, sce_imubac_disc$CellType != ""]

data_logit <- computeFeatures(sce = sce_imubac_disc, featureType = 'prop', 
                              cellTypeCol = 'CellType', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])

Age <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"])
GenderM <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")])

clinicaldata_info <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_imubac_disc, featureType = 'mean', 
                              cellTypeCol = 'CellType', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers, assay = "normexprs")
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))
```


## Compute features on study 3
## Proportions
```{r}
sce_imubac_val <- sce_imubac_val[, sce_imubac_val$mg_clusters != "other"]
sce_imubac_val <- sce_imubac_val[, sce_imubac_val$mg_clusters != ""]
sce_imubac_val$mg_clusters <- droplevels(sce_imubac_val$mg_clusters)

data_val_logit <- computeFeatures(sce = sce_imubac_val, featureType = 'prop', 
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_val <- rownames(data_val_logit)
condition_val <- factor(clinicaldata_val[match(row_names_val,clinicaldata_val$sample_id),
                                      "Gensini_bin"])
colnames(data_val_logit) <- as.factor(colnames(data_val_logit))

Age <- clinicaldata_val[match(row_names_val,clinicaldata_val$sample_id),
                                      "Age"]
GenderM <- clinicaldata_val[match(row_names_val,clinicaldata_val$sample_id),
                                      c("GenderM")]

clinicaldata_info_val <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType_val <- computeFeatures(sce = sce_imubac_val, featureType = 'mean', 
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers, assay = "normexprs")
```


# Generate Tree
```{r, eval=T}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_logit')
    `14+ monos` <- myeloid$AddChild('14+ monos_logit')
    `16+ monos` <- myeloid$AddChild('16+ monos_logit') 
    pDCs <- myeloid$AddChild('pDCs_logit')
    `CD141+ DCs` <- myeloid$AddChild('CD141+ DCs_logit')
    `CD1c+ DCs` <- myeloid$AddChild('CD1c+ DCs_logit')
  NK <- root$AddChild('NK_logit')
  `B cells` <- root$AddChild('B cells_logit')
  CD3 <- root$AddChild('CD3_logit')
    CD8hi <- CD3$AddChild('CD8hi_logit')
    CD8lo <- CD3$AddChild('CD8lo_logit')
    CD4 <- CD3$AddChild('CD4_logit')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_logit')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_logit')
  # other <- root$AddChild('other_logit')
```

# Generate grouping structures
```{r}
nclust <- ncol(data_logit)
# root <- as.phylo(root)
order <- as.phylo(root)$tip.label

tree <- treekoR:::findChildren(ggtree(as.phylo(root),
                                            ladderize = F, layout = 'dendrogram'))

groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
# X_train <- cbind(X_train, clinicaldata_info)
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_val_logit, markerMeanCellType_val) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
# X_test <- cbind(X_test, clinicaldata_info_3)
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```
## Predict on test data
```{r}
y_test <- as.numeric(levels(condition_val))[condition_val]
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../dioscRi_manuscript/auc_data//mg_iMUBAC_auc_all_study_3.csv')

```

# Visualise Resulting Tree
```{r, fig.height=20, fig.width=40}
p.all <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG",
                   trainingData = X_train, title = "Manual Gating")

# ggsave(file="../Plots/fs_tree.svg", plot=p, width=18, height=16, dpi = 1200)
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit %>%
  as.matrix()
```

## Testing
```{r}
X_test <- data_val_logit %>%
  as.matrix()

```

```{r}
groups <- generateGroups(tree = tree, nClust = 11, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```
## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../dioscRi_manuscript/auc_data/mg_iMUBAC_auc_prop_study_3.csv')

```

## Look at tree plot
```{r, fig.height=20, fig.width=40}
p.prop <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", 
                   trainingData = X_train, heatmap = FALSE)
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType %>%
  as.matrix()
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_val %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../dioscRi_manuscript/auc_data/mg_iMUBAC_auc_means_study_3.csv')

```

```{r, fig.height=24, fig.width=40}
order <- gsub("_logit", "", order)
order <- gsub("_", " ", order)
p.means <- plotHeatmap(fit$fit, type = "MG", order = order)
```

```{r, fig.width=20, fig.height=12, eval=TRUE}
pdf(file = "../dioscRi_manuscript/Plots/feature_trees_mg_iMUBAC.pdf",   # The directory you want to save the file in
    width = 20, # The width of the plot in inches
    height = 12) # The height of the plot in inches
p.prop_mean <- cowplot::plot_grid(p.prop, p.means, nrow = 2, 
                                  labels = LETTERS[1:2], align = "hv")
cowplot::plot_grid(p.prop_mean, p.all, ncol = 2, labels = c("", "C"))
dev.off()
```

