---
title: "Bioheart Prediction"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
library(dioscRi)
library(ggplot2)
library(FuseSOM)
library(SingleCellExperiment)
library(tidyverse)
library(data.table)
library(data.tree)
tensorflow::set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')
```
# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
cluster_col <- "clusters_11"
```


# Load normalised data
```{r}
load('../../../sce_dat/study_4_MMD_VAE_updated.RData')
load('../../../sce_dat/study_3_MMD_VAE_updated.RData')
```

## Use FuseSOM to cluster the normalised data
```{r, eval=T}
# run FuseSOM clustering
nclust = 11
sce_norm$cluster_col <- NULL
sce_norm <- runFuseSOM(sce_norm, numClusters = nclust, assay = 'norm',
                       verbose = FALSE, clusterCol = 'clusters_norm')
# sce_norm$clusters <- sce_temp$clusters
```

# Annotate clustering with Helen's annotations
```{r, eval=F}
# sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
# sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != ""]

sce_norm$clusters_norm <- recode_factor(sce_norm$clusters_norm,
                                 "cluster_1" = "NK", # NK Cells
                                 "cluster_2" = "Not Naive CD8+ T cells", # Not Naive CD8+ T cells
                                 "cluster_3" = "Naive CD8+ T cells", # Naive CD8+ T cells
                                 "cluster_4" = "NK/Mixed", # NK Cells
                                 "cluster_5" = "Naive B Cells", # mix of monocytes and dendritic cells
                                 "cluster_6" = "CD4+ Tconv", # mix of naïve CD4+ T cells, B cells and NK cells
                                 "cluster_7" = "Memory B cells", # B cells
                                 "cluster_8" = "16+ monocytes", # NK cells
                                 "cluster_9" = "CD141+ DC/pDCs", # central memory CD8+ T cells
                                 "cluster_10" = "14+ monocytes", # naïve CD8+ T cells
                                 "cluster_11" = "CD4+ Treg" # terminal effector memory revertant CD8+ T cells
                                 )
```

# Look at the clustering performance per sample
```{r}
df_norm <- assay(sce_norm, type = "norm") %>%
  t() %>%
  as.data.frame() %>%
  # mutate(cellTypes = sce_norm$mg_cell_type_distinct) %>%
  mutate(predicted_norm = sce_norm$clusters_11)
  # mutate(predicted_raw = sce_norm$clusters_raw) %>%
  # mutate(sample_id = sce_norm$sample_id)
```


# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_norm, type = "VAE") %>%
  dplyr::mutate(cellTypes = as.factor(sce_norm$clusters_11))

test_x <- reducedDim(sce_3_norm, type = "VAE")
```

```{r, eval=T}
df_3_clusters <- trainCellTypeClassifier(trainX = train_x, testX = test_x, model = 'lda')

table(df_3_clusters)

sce_3_norm$clusters <- as.factor(df_3_clusters)
```


# Compute training features
## Proportions
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)
# 
# row_names <- clinicaldata$sample_id
# 
# Age <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       "Age"]
# GenderM <- clinicaldata[match(row_names,clinicaldata$sample_id),
#                                       c("GenderM")]
# 
# clinicaldata_info <- cbind(Age, GenderM)
```

## Generate features
### Proportions
```{r}
clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
# clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)
clinicaldata_3$Gensini_bin <- as.factor(clinicaldata_3$Gensini_bin)
```


# Compute training features
## Proportions
```{r}
data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = F, useMarkers = useMarkers)

row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])
# colnames(data_logit) <- as.factor(colnames(data_logit))

Age <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"])
GenderM <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")])

clinicaldata_info <- cbind(Age, GenderM)

# data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
#                               cellTypeCol = 'cluster_col', sampleCol = 'sample_id', 
#                               logit = F, useMarkers = useMarkers)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', 
                              cellTypeCol = cluster_col, sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))

# markerMeanCellType_rsc <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = 'norm_rescaled',
#                               cellTypeCol = 'cluster_col', sampleCol = 'sample_id', 
#                               logit = T, useMarkers = useMarkers)
# colnames(markerMeanCellType_rsc) <- as.factor(colnames(markerMeanCellType_rsc))
```


## Compute features on study 3
## Proportions
```{r}
data_3_logit <- computeFeatures(sce = sce_3_norm, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_3 <- rownames(data_3_logit)
condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])
colnames(data_3_logit) <- as.factor(colnames(data_3_logit))

Age <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType_3 <- computeFeatures(sce = sce_3_norm, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))
```


# Generate grouping structures
```{r}
nclust <- length(unique(sce_norm$clusters_norm))
trees <- generateTree(features = data_logit, method = "ward")
tree <- trees$tree
order <- trees$order
groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Add the clinical data in
```{r, eval=F}
groups[[length(groups) + 1]] <- 309:310
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
# X_train <- cbind(X_train, clinicaldata_info)
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_3_logit, markerMeanCellType_3) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
# X_test <- cbind(X_test, clinicaldata_info_3)
y_test <- as.numeric(levels(condition_3))[condition_3]
```


## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, modelSummary = F,
                groups = groups, penalty = "grLasso", seed = 1994)
```
## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = condition_3, title = "Study 3 - All AUC =")
test_auc$plot
```

# Visualise Resulting Tree
```{r, fig.height=10, fig.width=12}
visualiseModelTree(fit = fit$fit, tree = tree,
                   trainingData = X_train, title = "Unsupervised Clustering")
```

```{r, eval=T, fig.height=10, fig.width=16}
# CCR2 (OR 1.12), CCR4 (OR 1.08), CD38 and CD45RO (OR 1.13), HLA-DR (OR 1.06), and Ki67 (1.22)
stats_mean <- getSigFeatures(fit$fit, type = 'mean',  
                         mean = markerMeanCellType, clinicalData = clinicaldata, 
                         outcome = "Gensini_bin")

sigFeatures <- stats_mean$sigFeatures
stats <- stats_mean$stats

# save(stats_mean, file = "../../../discovery_cohort_sig_stats_unsupervised.rda")


plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin")

plotSigFeatures(stats = stats, sigFeatures = sigFeatures, outcome = "Gensini bin", type = "boxplots")
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit %>%
  as.matrix()
```

## Testing
```{r}
X_test <- data_3_logit %>%
  as.matrix()

```

```{r}
groups <- generateGroups(tree = tree, nClust = 11, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train,
                groups = groups, penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot
```
## Look at tree plot
```{r, fig.height=5, fig.width=6}
visualiseModelTree(fit = fit$fit, tree = tree, 
                   trainingData = X_train, heatmap = FALSE)
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType %>%
  as.matrix()
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_3 %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train,
                groups = groups, penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot

```


```{r, fig.height=5, fig.width=6}
plotHeatmap(fit$fit, order = order, markerOrder = )
```



