---
title: "BioHEART Prediction cyCombine"
author: "Elijah Willie"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
library(dioscRi)
library(ggplot2)
library(FuseSOM)
library(SingleCellExperiment)
library(tidyverse)
library(data.table)
library(data.tree)
tensorflow::set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')
```
\
# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
```

# Load the cyCombine normalised data
```{r}
load('../../../sce_dat/bioheart_cyCombine_batch_corrected.RData')
```

## Subset the data
```{r, eval=T}
sce_cycombine_disc <- sce_cycombine[, sce_cycombine$batch == "4"]


sce_cycombine_val <- sce_cycombine[, sce_cycombine$batch == "3"]

```

## Use FuseSOM to cluster the normalised data
```{r, eval=T}
# run FuseSOM clustering
nclust = 11
sce_cycombine_disc$clusters_norm <- NULL
sce_cycombine_disc <- runFuseSOM(sce_cycombine_disc, numClusters = nclust, 
                                 assay = 'corrected',
                       verbose = FALSE, clusterCol = 'clusters_norm')
```
# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- assay(sce_cycombine_disc, "corrected") %>%
    t() %>%
    as.data.frame() %>%
  mutate(cellTypes = as.factor(sce_cycombine_disc$clusters_norm))

test_x <- assay(sce_cycombine_val, "corrected") %>%
    t() %>%
    as.data.frame()
```

```{r, eval=T}
val_clusters <- trainCellTypeClassifier(trainX = train_x, 
                                        testX = test_x, model = 'lda')

table(val_clusters)

sce_cycombine_val$clusters_norm <- as.factor(val_clusters)
```

# Compute training features
## Proportions
```{r}
clinicaldata <- colData(sce_cycombine_disc) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)
```

## Generate features
### Proportions
```{r}
clinicaldata_val <- colData(sce_cycombine_val) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
```

# Compute training features
## Proportions
```{r}
data_logit <- computeFeatures(sce = sce_cycombine_disc, featureType = 'prop', 
                              cellTypeCol = 'clusters_norm', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_cycombine_disc, featureType = 'mean', 
                              cellTypeCol = 'clusters_norm', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers, assay = "corrected")
colnames(markerMeanCellType) <- as.factor(colnames(markerMeanCellType))
```


## Compute features on study 3
## Proportions
```{r}
data_val_logit <- computeFeatures(sce = sce_cycombine_val, featureType = 'prop', 
                              cellTypeCol = 'clusters_norm', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_val <- rownames(data_val_logit)
condition_val <- factor(clinicaldata_val[match(row_names_val,clinicaldata_val$sample_id),
                                      "Gensini_bin"])
colnames(data_val_logit) <- as.factor(colnames(data_val_logit))
```

## Means
```{r, eval=T}
markerMeanCellType_val <- computeFeatures(sce = sce_cycombine_val, featureType = 'mean', 
                              cellTypeCol = 'clusters_norm', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers, assay = "corrected")
```


# Generate grouping structures
```{r}
nclust <- 11
trees <- generateTree(features = data_logit, method = "ward")
tree <- trees$tree
order <- trees$order
groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()

# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_val_logit, markerMeanCellType_val) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
y_test <- as.numeric(levels(condition_val))[condition_val]
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 - All AUC =")
test_auc$plot
```

# Visualise Resulting Tree
```{r, fig.height=10, fig.width=12}
visualiseModelTree(fit = fit$fit, tree = tree,
                   trainingData = X_train, title = "Unsupervised Clustering")
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit %>%
  as.matrix()
```

## Testing
```{r}
X_test <- data_val_logit %>%
  as.matrix()

```

```{r}
groups <- generateGroups(tree = tree, nClust = 11, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
groups <- lapply(groups, sort)
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot
```

## Look at tree plot
```{r, fig.height=5, fig.width=6}
visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", 
                   trainingData = X_train, heatmap = FALSE)
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType %>%
  as.matrix()
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_val %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, 
                    title = "Study 3 - All AUC =")
test_auc$plot
```

```{r, fig.height=5, fig.width=6}
plotHeatmap(fit$fit, order = order)
```

