---
title: "Code to generate various parts of Figure 3"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tensorflow)
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(data.table)
  library(SingleCellExperiment)
  library(ROCR)
  library(pROC)
  library(data.tree)
  library(ggtree)
  library(ape)
  library(treekoR)
  library(ggiraph)
  library(scales)
  library(ggnewscale)
  library(dioscRi)
})
source('../scripts/normalisation_plotting.R')
set_random_seed(1994)
```

# Set the markers
```{r}
useMarkers <- c('HLA-DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c-PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Load normalised data
```{r}
load('../../sce_dat/study_4_MMD_VAE_updated.RData')
load('../../sce_dat/study_3_MMD_VAE_updated.RData')
```

# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_norm, type = "VAE") %>%
  dplyr::mutate(cellTypes = sce_norm$mg_cell_type_distinct)


test_x <- reducedDim(sce_3_norm, type = "VAE") %>%
  as.data.frame()
```

## Fit SVM model
```{r, eval=T}
df_3_clusters <- dioscRi::trainCellTypeClassifier(trainX = train_x, testX = test_x, 
                                                  model = 'lda')

table(df_3_clusters)

sce_3_norm$mg_clusters <- df_3_clusters
```

# Setup Clinical information
## Study 4
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini, Gender, Age) %>%
  distinct()
clinicaldata$Gensini_bin <- factor(if_else(clinicaldata$Gensini > 0, 1, 0))

encoded_gender <- model.matrix(~Gender-1, data=clinicaldata)
clinicaldata <- cbind(clinicaldata, encoded_gender)

```

## Study 3
```{r}
clinicaldata_3 <- colData(sce_3_norm) %>%
  as.data.frame() %>%
  dplyr::select(sample_id, Gensini_bin, Gender, Age) %>%
  distinct()
# clinicaldata_3$Gensini_bin <- factor(if_else(clinicaldata_3$Gensini > 0, 1, 0))
encoded_gender_3 <- model.matrix(~Gender-1, data=clinicaldata_3)
clinicaldata_3 <- cbind(clinicaldata_3, encoded_gender_3)
clinicaldata_3$Gensini_bin <- as.factor(clinicaldata_3$Gensini_bin)
```


# Compute training features
## Proportions
```{r}
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != "other"]
sce_norm <- sce_norm[, sce_norm$mg_cell_type_distinct != ""]
# sce_norm$mg_cell_type_distinct <- droplevels(sce_norm$mg_cell_type_distinct)
data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', 
                              sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Gensini_bin"])

data <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = F, useMarkers = useMarkers)
Age <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "Age"])
GenderM <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      c("GenderM")])

clinicaldata_info <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', assay = "norm",
                              cellTypeCol = 'mg_cell_type_distinct', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```


## Compute features on study 3
## Proportions
```{r}
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != "other"]
sce_3_norm <- sce_3_norm[, sce_3_norm$mg_clusters != ""]
sce_3_norm$mg_clusters <- droplevels(sce_3_norm$mg_clusters)

data_3_logit <- computeFeatures(sce = sce_3_norm, featureType = 'prop', 
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_3 <- rownames(data_3_logit)
condition_3 <- factor(clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Gensini_bin"])

Age <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      "Age"]
GenderM <- clinicaldata_3[match(row_names_3,clinicaldata_3$sample_id),
                                      c("GenderM")]

clinicaldata_info_3 <- cbind(Age, GenderM)
```

## Means
```{r, eval=T}
markerMeanCellType_3 <- computeFeatures(sce = sce_3_norm, featureType = 'mean',  assay = "norm",
                              cellTypeCol = 'mg_clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

# Generate Tree
```{r, eval=T}
root <- Node$new("Root")
  myeloid <- root$AddChild('Myeloid_logit')
    `14+ monos` <- myeloid$AddChild('14+ monos_logit')
    `16+ monos` <- myeloid$AddChild('16+ monos_logit') 
    pDCs <- myeloid$AddChild('pDCs_logit')
    `CD141+ DCs` <- myeloid$AddChild('CD141+ DCs_logit')
    `CD1c+ DCs` <- myeloid$AddChild('CD1c+ DCs_logit')
  NK <- root$AddChild('NK_logit')
  `B cells` <- root$AddChild('B cells_logit')
  CD3 <- root$AddChild('CD3_logit')
    CD8hi <- CD3$AddChild('CD8hi_logit')
    CD8lo <- CD3$AddChild('CD8lo_logit')
    CD4 <- CD3$AddChild('CD4_logit')
      `CD4+ Tconv` <- CD4$AddChild('CD4+ Tconv_logit')
      `CD4+ Treg` <- CD4$AddChild('CD4+ Treg_logit')
  # other <- root$AddChild('other_logit')
```

# Generate grouping structures
```{r}
nclust <- ncol(data_logit)
# root <- as.phylo(root)
order <- as.phylo(root)$tip.label
tree <- treekoR:::findChildren(ggtree(as.phylo(root),
                                            ladderize = F, layout = 'dendrogram'))

groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Add the clinical data in
```{r, eval=F}
groups[[length(groups) + 1]] <- 309:310
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
  # Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
# X_train <- cbind(X_train, clinicaldata_info)
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_3_logit, markerMeanCellType_3) %>%
  as.matrix()
# colnames(X_test_study_3) <- janitor::make_clean_names(colnames(X_test_study_3))
X_test <- predict(scaleVals, X_test)
# X_test <- cbind(X_test, clinicaldata_info_3)
y_test <- as.numeric(levels(condition_3))[condition_3]
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994, alpha = NULL)
```

## Predict on the training data
```{r}
train_auc <- plotAUC(fit = fit$fit, xTest = X_train, yTest = y_train, title = "Study 4 MG - All AUC =")
train_auc$plot
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../auc_data/mg_dioscRi_auc_all_study_3.csv')

```



# Visualise Resulting Tree (Figure 3B)
```{r, fig.height=20, fig.width=40}
# tiff("../Plots/test.png", units="in", width=16, height=10, res=300)
p.all <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", trainingData = X_train,
                   nodesToRemove =  c("Root", "Myeloid", "CD3", "CD4"), title = "Proportions + Marker Means")

pdf(file = "../Plots/Figure_3_tree_full_manual_gating.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 10) # The height of the plot in inches
    p.all 
dev.off()
# dev.off()
# ggsave(file="../Plots/mg_tree.svg", plot=p, width=18, height=16, dpi = 1200)
```

# Look at just the proportions
## Training
```{r}
# Combine all input data matrices
X_train <- data_logit
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test<- data_3_logit %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- generateGroups(tree = tree, nClust = nclust, proportions = data_logit,
                         means = markerMeanCellType, type = "prop")
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot
# 
write.csv(test_auc$preds,
          '../auc_data//mg_dioscRi_auc_prop_study_3.csv')

```

## Look at tree plot (Figure 3A)
```{r, fig.height=15, fig.width=40}
p.prop <- visualiseModelTree(fit = fit$fit, tree = tree, type = "MG", trainingData = X_train,
                   nodesToRemove =  c("Root", "Myeloid", "CD3", "CD4"), heatmap = FALSE)

pdf(file = "../Plots/Figure_3_tree_prop_manual_gating.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 10) # The height of the plot in inches
    p.prop 
dev.off()
```

# Look at just the means
## Training
```{r}
# Combine all input data matrices
X_train <- markerMeanCellType
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
```

## Testing
```{r}
X_test <- markerMeanCellType_3 %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```

```{r}
groups <- as.list(seq (1,ncol(markerMeanCellType),1)) 
```

## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train, groups = groups, 
                penalty = "grLasso", seed = 1994)
```

## Predict on test data
```{r}
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Study 3 MG - All AUC =")
test_auc$plot

write.csv(test_auc$preds,
          '../auc_data/mg_dioscRi_auc_means_study_3.csv')



```

## Look at heatmap plot (Figure 3C)
```{r, fig.height=15, fig.width=40}
order <- gsub("_logit", "", order)
order <- gsub("_", " ", order)
p.means <- plotHeatmap(fit$fit, type = "MG", order = order)

pdf(file = "../Plots/Figure_3_tree_means_manual_gating.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 8) # The height of the plot in inches
    p.means 
dev.off()
```

# Define some color schemes
```{r}
all_colors <- c(
  "#1f77b4",  # Blue
  "#ff7f0e",  # Orange
  "#9467bd",  # Purple
  "#8c564b",  # Brown
  "#e377c2",  # Pink
  "#7f7f7f",  # Gray
  "#bcbd22",  # Olive-yellow
  "#17becf",  # Cyan
  "#aec7e8",  # Light blue
  "#c5b0d5"   # Lavender
)
my_colors_1 <- c(
  "Prop" = all_colors[[3]],         # soft blue
  "Means" = all_colors[[4]],        # warm amber/orange
  "Means + Prop" = all_colors[[5]]  # muted teal (not green-looking)
)
my_colors_2 <- c(
  "Age + Sex" = all_colors[[8]],         # soft blue
  "Means + Prop" = all_colors[[9]],        # warm amber/orange
  "Means + Prop + Age + Sex" = all_colors[[10]]  # muted teal (not green-looking)
)
```


## Figure 3D
```{r}
prop <- fread('../auc_data/fs_dioscRi_auc_prop_study_3.csv')
means <- fread('../auc_data/fs_dioscRi_auc_means_study_3.csv')
full <- fread('../auc_data/fs_dioscRi_auc_all_study_3.csv')

prop_roc <- roc(prop, Truth, Predicted)
means_roc <- roc(means, Truth, Predicted)
full_roc <- roc(full, Truth, Predicted)

p.fs <- ggrocs(rocs = list("Prop" = prop_roc, 
                   "Means" = means_roc,
                   "Means + Prop" = full_roc), 
        data_name = "Performance for unsupervised clustering", legend_title = "Data",
        legend_order = c("Prop","Means", "Means + Prop"), 
        color_scale = my_colors_1, which_plot = "bar")

pdf(file = "../Plots/Figure_3_unsupervised_bar_auc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.fs
dev.off()
```
## Figure 3E
```{r}
prop_mg <- fread('../auc_data/mg_dioscRi_auc_prop_study_3.csv')
means_mg <- fread('../auc_data/mg_dioscRi_auc_means_study_3.csv')
full_mg <- fread('../auc_data/mg_dioscRi_auc_all_study_3.csv')

prop_mg_roc <- roc(prop_mg, Truth, Predicted)
means_mg_roc <- roc(means_mg, Truth, Predicted)
full_mg_roc <- roc(full_mg, Truth, Predicted)

p.mg <- ggrocs(rocs = list("Prop" = prop_mg_roc, 
                   "Means" = means_mg_roc,
                   "Means + Prop" = full_mg_roc), 
        data_name = "Performance for manual gated clusters", legend_title = "Data",
        legend_order = c("Prop","Means", "Means + Prop"),
        color_scale = my_colors_1, which_plot = "bar")

pdf(file = "../Plots/Figure_3_manual_gating_bar_auc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.mg 
dev.off()
```
## Figure 3F
```{r}
age_sex <- fread('../auc_data/fs_glasso_auc_all_study_3_clinical_only.csv')
full_fs <- full_fs <- fread('../auc_data/fs_dioscRi_auc_all_study_3.csv')
full_age_sex <- fread('../auc_data/fs_dioscRi_auc_all_clinical_study_3.csv')

age_sex_roc <- roc(age_sex, Truth, Predicted)
full_fs_roc <- roc(full_fs, Truth, Predicted)
full_age_sex_roc <- roc(full_age_sex, Truth, Predicted)

p.fs.clinical <- ggrocs(rocs = list("Age + Sex" = age_sex_roc, 
                   "Means + Prop" = full_fs_roc,
                   "Means + Prop + Age + Sex" = full_age_sex_roc), 
        data_name = "Performance for unsupervised clustering", legend_title = "Data",
        color_scale = my_colors_2)

pdf(file = "../Plots/Figure_3_unsupervised_clinical_bar_auc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.fs.clinical
dev.off()
```

## Figure 3G
```{r}
age_sex <- fread('../auc_data/fs_glasso_auc_all_study_3_clinical_only.csv')
full_mg <- full_fs <- fread('../auc_data/mg_dioscRi_auc_all_study_3.csv')
full_mg_age_sex <- fread('../auc_data/mg_dioscRi_auc_all_clinical_study_3.csv')

age_sex_roc <- roc(age_sex, Truth, Predicted)
full_mg_roc <- roc(full_mg, Truth, Predicted)
full_mg_age_sex_roc <- roc(full_mg_age_sex, Truth, Predicted)

p.mg.clinical <- ggrocs(rocs = list("Age + Sex" = age_sex_roc, 
                   "Means + Prop" = full_mg_roc,
                   "Means + Prop + Age + Sex" = full_mg_age_sex_roc), 
        data_name = "Performance for manual gated clusters", legend_title = "Data",
        color_scale = my_colors_2)

pdf(file = "../Plots/Figure_3_manual_gating_clinical_bar_auc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.mg.clinical
dev.off()
```