---
title: "Code to generate various parts of Figure 2"
author: "Elijah WIllie"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(data.table)
  library(SingleCellExperiment)
  library(ROCR)
  library(scales)
  library(ggnewscale)
  library(cvAUC)
  library(pROC)
  library(Polychrome)
  library(RColorBrewer)
})
source('~/Documents/PhD/bioheart_analysis/dioscRi_manuscript/scripts/normalisation_plotting.R')
set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Load in sce objects
```{r}
load('../../sce_dat/study_4_MMD_VAE_final.RData')
load('../../sce_dat/study_3_MMD_VAE_final.RData')
load('../../sce_dat/bioheart_cyCombine_batch_corrected.RData')
load('../../sce_dat/bioheart_iMUBAC_batch_corrected.RData')
```

# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
```

# Define cell types of interest
```{r}
# only use celltypes that is present in both batches
final_celltypes <- c(
   "B cells",
   "CD8hi",
   "14+ monos",
   "CD4+ Tconv",
   "CD8lo",
   "NK",
   "16+ monos",
   "CD4+ Treg",
   "pDCs",
   "CD141+ DCs",
   "CD1c+ DCs"
    )
```


# Load in teh raw data for the BioHEART datasets
```{r}
df_4_raw <- assay(sce_norm, 'raw') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_norm$mg_cell_type_distinct) %>%
    dplyr::mutate(batch = sce_norm$CyTOF.Batch) %>%
    dplyr::mutate(sample_id = sce_norm$sample_id) %>%
    dplyr::mutate(Age = sce_norm$Age) %>%
    dplyr::mutate(Gender = sce_norm$Gender) %>%
    dplyr::mutate(Statin = sce_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_norm$gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_3_raw <- assay(sce_3_norm, 'raw') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_3_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_3_norm$CellTypes) %>%
    dplyr::mutate(batch = sce_3_norm$Batch) %>%
    dplyr::mutate(sample_id = sce_3_norm$sample_id) %>%
    dplyr::mutate(Age = sce_3_norm$Age) %>%
    dplyr::mutate(Gender = sce_3_norm$Gender) %>%
    dplyr::mutate(Statin = sce_3_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_3_norm$Gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_raw <- rbind(df_3_raw, df_4_raw)

```


# Load in the dioscRi normalised data for the BioHEART datasets
```{r}
df_4_norm <- assay(sce_norm, 'norm') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_norm$mg_cell_type_distinct) %>%
    dplyr::mutate(batch = sce_norm$CyTOF.Batch) %>%
    dplyr::mutate(sample_id = sce_norm$sample_id) %>%
    dplyr::mutate(Age = sce_norm$Age) %>%
    dplyr::mutate(Gender = sce_norm$Gender) %>%
    dplyr::mutate(Statin = sce_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_norm$gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_3_norm <- assay(sce_3_norm, 'norm') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_3_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_3_norm$CellTypes) %>%
    dplyr::mutate(batch = sce_3_norm$Batch) %>%
    dplyr::mutate(sample_id = sce_3_norm$sample_id) %>%
    dplyr::mutate(Age = sce_3_norm$Age) %>%
    dplyr::mutate(Gender = sce_3_norm$Gender) %>%
    dplyr::mutate(Statin = sce_3_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_3_norm$Gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_norm <- rbind(df_3_norm, df_4_norm)
```

# Load in the cyCombine normalised data for the BioHEART datasets
```{r}
df_norm_cycombine <- assay(sce_cycombine, 'corrected') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype = sce_cycombine$CellType) %>%
    dplyr::mutate(sample_id = sce_cycombine$sample_id) %>%
    dplyr::mutate(batch = sce_cycombine$batch) %>%
    dplyr::mutate(Gensini = sce_cycombine$Gensini_bin) %>%
    dplyr::mutate(Statin = sce_cycombine$Statin) %>%
    dplyr::mutate(Age = sce_cycombine$Age) %>%
    dplyr::mutate(Gender = sce_cycombine$Gender) %>%
    # dplyr::filter(batch == 4)  %>%
    dplyr::filter(Celltype %in% final_celltypes)

```

# Load in the iMUBAC normalised data for the BioHEART datasets
```{r}
df_norm_imubac <- assay(sce_imubac, 'normexprs') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype = sce_imubac$CellType) %>%
    dplyr::mutate(sample_id = sce_imubac$sample_id) %>%
    dplyr::mutate(batch = sce_imubac$batch) %>%
    dplyr::mutate(Gensini = sce_imubac$Gensini_bin) %>%
    dplyr::mutate(Statin = sce_imubac$Statin) %>%
    dplyr::mutate(Age = sce_imubac$Age) %>%
    dplyr::mutate(Gender = sce_imubac$Gender) %>%
    # dplyr::filter(batch == 4) %>%
    dplyr::filter(Celltype %in% final_celltypes)
```

# Cluster the discovery 
```{r}
# get the number of clusters
numClusters = length(final_celltypes)
    
# Run FuseSOM for each dataset
clusters <- runFuseSOM(rbind(df_norm[, useMarkers]), 
                                    numClusters = numClusters)$clusters
gc()
    
clusters_cycombine <- runFuseSOM(df_norm_cycombine[, useMarkers],
                                    numClusters = numClusters)$clusters
gc()
    
clusters_imubac <- runFuseSOM(df_norm_imubac[, useMarkers],
                                    numClusters = numClusters)$clusters
gc()
    
clusters_raw <- runFuseSOM(df_raw[, useMarkers], 
                                    numClusters = numClusters)$clusters
gc()
```

# Compute ARI with cell type
```{r}
# record the scores for manual gates
ari_dscri_mg <- aricode::ARI(clusters, df_norm$Celltype)
ari_cycombine_mg <- aricode::ARI(clusters_cycombine, df_norm_cycombine$Celltype)
ari_imubac_mg <- aricode::ARI(clusters_imubac, df_norm_imubac$Celltype)
ari_raw_mg <- aricode::ARI(clusters_raw, df_raw$Celltype)

mg_ari <- c(ari_dscri_mg, ari_cycombine_mg, ari_imubac_mg, ari_raw_mg)
```

# generate Figure 2I
```{r}
ari_df <- data.frame(
                     dioscRi = ari_dscri_mg,
                     cyCombine = ari_cycombine_mg,
                     iMUBAC = ari_imubac_mg,
                     Raw = ari_raw_mg
                     ) %>%
    reshape2::melt() %>%
    dplyr::rename(Comparison = variable,
           ARI = value) %>%
    dplyr::mutate(Comparison = factor(Comparison, 
                                      levels = c("dioscRi","cyCombine","iMUBAC","Raw")))

p_ari <- ari_df %>%
  ggplot(aes(Comparison, ARI, fill = Comparison)) +
  geom_col(width = 0.4) +  # ← adjust this width to control bar thickness
  ggtitle(label = "Combined discovery and validation") +
    theme_bw() +
  theme(
    axis.text.x = element_text(size = 12, hjust = 1),
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_blank(), 
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.position = "none",  # lowercase 'none'
    strip.text.x = element_text(size = 12, colour = "black", face = "bold"),
    aspect.ratio = 1
  ) +
  scale_fill_npg()

pdf(file = "../Plots/Figure_2_ARI.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p_ari
dev.off()

```



# Generate figures 2E and 2F
## dioscRi discovery
```{r}
# Compute density and find minimum
valley_lines <- df_4_norm %>%
  dplyr::filter(sample_id %in% c("87", "88")) %>%
  dplyr::select(CD3, sample_id) %>%
  group_by(sample_id) %>%
  summarise(
    min_x = {
      d <- density(CD3, na.rm = TRUE, from = 0, to = 0.75)
      d$x[which.min(d$y)]
    },
    .groups = "drop"
  )

# Plot with density and per-sample valley line
p_post <- df_4_norm %>%
  filter(sample_id %in% c("87", "88")) %>%
  mutate(Sample = factor(sample_id)) %>%
  ggplot(aes(CD3, color = Sample)) +
  geom_density() +
  scale_x_continuous(limits = c(0, 0.9), oob = scales::oob_keep) +
  geom_vline(
    data = valley_lines,
    aes(xintercept = min_x),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) +
  ggtitle("dioscRi") +
  ylab("Density") +
  xlab("") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    # legend.position = "none",
    aspect.ratio = 1,
    strip.text.x = element_text(size = 12, colour = "black", face = "bold")
  ) +
  scale_color_npg()

pdf(file = "../Plots/Figure_2_density_post.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p_post
dev.off()
```

## Raw discovery
```{r}
# put the data on the same scale as the normalised for comparison
preProcValues_raw <- preProcess(df_4_raw[, useMarkers], method = c("range"))
df_4_raw[, useMarkers] <- predict(preProcValues_raw, df_4_raw[, useMarkers])

p_pre <- df_4_raw  %>%
    dplyr::filter(sample_id %in% c("87", "88")) %>%
    dplyr::select(CD3, sample_id) %>%
    mutate(Sample = as.factor(sample_id)) %>%
    # reshape2::melt(id.vars = "sample_id") %>%
    # dplyr::mutate(Sample = as.factor(sample_id)) %>%
    ggplot(aes(CD3, color = Sample)) +
    # facet_wrap(~variable, scales = "free") +
    geom_density() +
    scale_x_continuous(limits = c(0,0.9), oob = scales::oob_keep) +
    geom_vline(
    data = valley_lines,
    aes(xintercept = min_x),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) +
    ggtitle("Raw") +
    ylab("Density") + 
    xlab("") +
    theme_bw() +
        theme(
      axis.text.x = element_text(size = 12),
      plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 12),
      axis.title.x = element_text(size = 12),
      aspect.ratio = 1,
      # legend.position = "none",
      strip.text.x = element_text(size = 12, 
                                  colour = "black", face = 'bold')
    ) +
    scale_color_npg()

pdf(file = "../Plots/Figure_2_density_pre.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p_pre
dev.off()
```


# Bi-axial plots of normalised expression for Figures 2A, 2B, 2C, and 2D
```{r}
# helper density function
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
```

## 2A
```{r}
p.raw <- df_4_raw %>%
        dplyr::filter(sample_id == 87) %>%
        dplyr::mutate(density = get_density(CD3, `HLA-DR`, n = 100)) %>%
        ggplot(aes(x = CD3, y = `HLA-DR`, color = density)) + 
    geom_point(size = 1) + 
    geom_vline(
    data = valley_lines,
    aes(xintercept = min_x - 0),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) + 
   scale_x_continuous(limits = c(0,0.9), oob = scales::oob_keep) +
    scale_y_continuous(limits = c(0,0.91), oob = scales::oob_keep) +
    scale_color_gradientn(trans = "sqrt", colors = c("navy", rev(brewer.pal(11, "Spectral")))) +
      theme_bw() +
        ggtitle(label = paste0("Raw - Sample 87")) +
        theme(
            legend.title = element_text(color = "black", size = 12),
            plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
            legend.text = element_text(color = "black", size = 12),
            legend.position = "none",
            axis.text.x = element_text(size = 12),
            axis.text.y = element_text(size = 12),
            aspect.ratio = 1
          )

pdf(file = "../Plots/Figure_2_87_raw.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.raw
dev.off()
```

# 2C
```{r}
p.norm <- df_4_norm %>%
        # dplyr::filter(batch == 4) %>%
        dplyr::filter(sample_id == 87) %>%
    dplyr::mutate(density = get_density(CD3, `HLA-DR`, n = 100)) %>%
        ggplot(aes(x = CD3, y = `HLA-DR`, colour = density)) + 
        geom_point(size = 1) +
        # geom_rect(aes(xmin = 0.4, xmax = 0.8, ymin = 0, ymax = 0.75), fill = NA, color = "green", linetype = "dashed") +
    geom_vline(
    data = valley_lines,
    aes(xintercept = min_x - 0),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) + 
       scale_x_continuous(limits = c(0,0.9), oob = scales::oob_keep) +
    scale_y_continuous(limits = c(0,0.91), oob = scales::oob_keep) +
    scale_color_gradientn(trans = "sqrt", colors = c("navy", rev(brewer.pal(11, "Spectral")))) +
      theme_bw() +
        ggtitle(label = paste0("dioscRi - Sample 87")) +
        theme(
            legend.title = element_text(color = "black", size = 12),
            plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
            legend.text = element_text(color = "black", size = 12),
            axis.text.x = element_text(size = 12),
            axis.text.y = element_text(size = 12),
            aspect.ratio = 1,
            legend.position = "none"
          )

pdf(file = "../Plots/Figure_2_87_norm.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.norm
dev.off()
```


## 2B
```{r}
p.raw.88 <- df_4_raw %>%
        # dplyr::filter(batch == 4) %>%
        dplyr::filter(sample_id == 88) %>%
    dplyr::mutate(density = get_density(CD3, `HLA-DR`, n = 100)) %>%
        ggplot(aes(x = CD3, y = `HLA-DR`, colour = density)) + 
        geom_point(size = 1) +
    # geom_rect(aes(xmin = 0.4, xmax = 0.8, ymin = 0, ymax = 0.75), fill = NA, color = "green", linetype = "dashed") +
    geom_vline(
    data = valley_lines,
    aes(xintercept = min_x - 0),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) + 
      scale_x_continuous(limits = c(0,0.9), oob = scales::oob_keep) +
    scale_y_continuous(limits = c(0,0.91), oob = scales::oob_keep) +
    scale_color_gradientn(trans = "sqrt", colors = c("navy", rev(brewer.pal(11, "Spectral")))) +
      theme_bw() +
        ggtitle(label = paste0("Raw - Sample 88")) +
        theme(
            legend.title = element_text(color = "black", size = 12),
            plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
            legend.text = element_text(color = "black", size = 12),
            legend.position = "none",
            axis.text.x = element_text(size = 12),
            axis.text.y = element_text(size = 12),
            aspect.ratio = 1
          )

pdf(file = "../Plots/Figure_2_88_raw.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.raw.88
dev.off()
```

# 2D
```{r}
p.norm.88 <- df_4_norm %>%
        # dplyr::filter(batch == 4) %>%
        dplyr::filter(sample_id == 88) %>%
    dplyr::mutate(density = get_density(CD3, `HLA-DR`, n = 100)) %>%
        ggplot(aes(x = CD3, y = `HLA-DR`, colour = density)) + 
        geom_point(size = 1) +
       # geom_rect(aes(xmin = 0.4, xmax = 0.8, ymin = 0, ymax = 0.75), fill = NA, color = "green", linetype = "dashed") +
    geom_vline(
    data = valley_lines,
    aes(xintercept = min_x - 0),
    linetype = "dashed",
    colour = "red",
    linewidth = 1
  ) + 
       scale_x_continuous(limits = c(0,0.9), oob = scales::oob_keep) +
    scale_y_continuous(limits = c(0,0.91), oob = scales::oob_keep) +
    scale_color_gradientn(trans = "sqrt", colors = c("navy", rev(brewer.pal(11, "Spectral")))) +
      theme_bw() +
        ggtitle(label = "dioscRi - Sample 88") +
        theme(
            legend.title = element_text(color = "black", size = 12),
            plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
            legend.text = element_text(color = "black", size = 12),
            axis.text.x = element_text(size = 12),
            axis.text.y = element_text(size = 12),
            aspect.ratio = 1,
            legend.position = "none"
          )

pdf(file = "../Plots/Figure_2_88_norm.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.norm.88
dev.off()
```



# Generate Figure 2H and 2G
```{r}
preProcValues_cycombine <- preProcess(df_norm_cycombine[, useMarkers], method = c("range"))
df_norm_cycombine[, useMarkers] <- predict(preProcValues_cycombine, df_norm_cycombine[, useMarkers])

preProcValues_imubac <- preProcess(df_norm_imubac[, useMarkers], method = c("range"))
df_norm_imubac[, useMarkers] <- predict(preProcValues_imubac, df_norm_imubac[, useMarkers])

preProcValues_raw <- preProcess(df_raw[, useMarkers], method = c("range"))
df_raw[, useMarkers] <- predict(preProcValues_raw, df_raw[, useMarkers])
```

## 2H
```{r}
# compute celltype summaries
celltypes <- final_celltypes

t_raw_mg <- get_summaries_per_celltype(data = df_raw, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_mg <- get_summaries_per_celltype(data = df_norm, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_cycombine_mg <- get_summaries_per_celltype(data = df_norm_cycombine, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_imubac_mg <- get_summaries_per_celltype(data = df_norm_imubac, 
                                   markers = useMarkers, celltypes = celltypes)

# plot boxplot
p.cv.box <- cvBoxPlotByMarker_ggplot(raw = t_raw_mg, dioscri = t_norm_mg, 
                                         cycombine = t_norm_cycombine_mg, 
                                         imubac = t_norm_imubac_mg,
                                         title = "Variance removed by normalisation")

pdf(file = "../Plots/Figure_2_relative_CV.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.cv.box
dev.off()
```

## 2G
### Normalised data
```{r}
k <- 27

df_norm_sub <- df_norm %>%
    dplyr::select(useMarkers)
# Step 2: Fit linear models and calculate R^2 for each PC
r_squared <- sapply(1:k, function(i) {
  model <- lm(df_norm_sub[, i] ~ Celltype, data = df_norm %>%
                  dplyr::select(Celltype, Gender, Gensini, Age))  # Fit linear model
  summary(model)$r.squared  # Extract R^2
})
```

### Raw data
```{r}

df_raw_sub <- df_raw %>%
    dplyr::select(useMarkers)
# Step 2: Fit linear models and calculate R^2 for each PC
r_squared_raw <- sapply(1:k, function(i) {
  model <- lm(df_raw_sub[, i] ~ Celltype, data = df_raw %>%
                  dplyr::select(Celltype, Gender, Gensini, Age))  # Fit linear model
  summary(model)$r.squared  # Extract R^2
})
```

### cyCombine
```{r}
df_norm_cycombine_sub <- df_norm_cycombine %>%
    dplyr::select(useMarkers)
# Step 2: Fit linear models and calculate R^2 for each PC
r_squared_cycombine <- sapply(1:k, function(i) {
  model <- lm(df_norm_cycombine_sub[, i] ~ Celltype, data = df_norm_cycombine %>%
                  dplyr::select(Celltype, Gender, Gensini, Age))  # Fit linear model
  summary(model)$r.squared  # Extract R^2
})

```

### iMUBAC
```{r}
df_norm_imubac_sub <- df_norm_imubac %>%
    dplyr::select(useMarkers)

# Step 2: Fit linear models and calculate R^2 for each PC
r_squared_imubac <- sapply(1:k, function(i) {
  model <- lm(df_norm_imubac_sub[, i] ~ Celltype, data = df_norm_imubac %>%
                  dplyr::select(Celltype, Gender, Gensini, Age))  # Fit linear model
  summary(model)$r.squared  # Extract R^2
})
```

## Compute the R2 values and generate plot
```{r}
rsq_scores_all <- data.frame(dioscRi = r_squared, cyCombine = r_squared_cycombine,
                            iMUBAC = r_squared_imubac, Raw = r_squared_raw,
                            Marker = useMarkers) %>%
    reshape2::melt(id.var = "Marker") %>%
    dplyr::rename(Method = variable) %>%
    dplyr::rename(Rsq = value) %>%
    dplyr::mutate(Method = factor(Method, 
                                         levels = c("dioscRi","cyCombine","iMUBAC","Raw")))

p.rsq_line <- rsq_scores_all %>%
    ggplot(aes(reorder(Marker, Rsq), Rsq, colour = Method)) +
    geom_line(aes(group = Method)) +
    geom_point(size = 3) +
    ggtitle("Variance explained by cell type") +
    xlab("") +
    ylab("R-squared") + 
    theme_bw() +
     theme(
   axis.text.y = element_text(size = 12),
                axis.text.x = element_text(size = 12, angle = 90, hjust = 1),
                axis.title.x = element_text(size = 12),
                axis.title.y = element_text(size = 12),
                plot.title = element_text(size = 12, hjust = 0.5, face = "bold"),
                legend.title = element_blank(),
                xis.line = element_line(colour = "black"),
   # aspect.ratio = 1,
    # legend.position = c(0.7, 0.25),
    strip.text.x = element_text(size = 12, colour = "black", face = 'bold'),
   plot.margin    = margin(t = 0, r = 0, b = 0, l = 0)
  ) + 
    scale_color_npg() 

pdf(file = "../Plots/Figure_2_rsq.pdf",   # The directory you want to save the file in
    width = 10, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.rsq_line
dev.off()
    
```

# Figure 2J and 2K
```{r}
# set colors of interest
all_colors <- c(
  "#1f77b4",  # Blue
  "#ff7f0e",  # Orange
  "#9467bd",  # Purple
  "#8c564b",  # Brown
  "#e377c2",  # Pink
  "#7f7f7f",  # Gray
  "#bcbd22",  # Olive-yellow
  "#17becf",  # Cyan
  "#aec7e8",  # Light blue
  "#c5b0d5"   # Lavender
)
my_colors_1 <- c("Raw" = all_colors[[1]], "Normalised" = all_colors[[2]])
```


## 2J
```{r, fig.height=6, fig.width=6}
baseline_fs <- fread('../auc_data/fs_dioscRi_auc_all_study_3_baseline.csv')
full_fs <- fread('../auc_data/fs_dioscRi_auc_all_study_3.csv')

baseline_fs_roc <- roc(baseline_fs, Truth, Predicted)
full_fs_roc <- roc(full_fs, Truth, Predicted)

p.fs.full <- ggrocs(rocs = list("Raw" = baseline_fs_roc, "Normalised" = full_fs_roc), 
        data_name = "Predicting coronoary artery disease", legend_title = "Unsupervised Clustering", 
        legend_order = c("Raw", "Normalised"), color_scale = my_colors_1, which_plot = "roc")

pdf(file = "../Plots/Figure_2_unsupervised_roc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.fs.full
dev.off()
```

## 2J
```{r, fig.height=6, fig.width=6}
baseline_mg <- fread('../auc_data/mg_dioscRi_auc_all_study_3_baseline.csv')
full_mg <- fread('../auc_data/mg_dioscRi_auc_all_study_3.csv')

full_mg_roc <- roc(full_mg, Truth, Predicted)
baseline_mg_roc <- roc(baseline_mg, Truth, Predicted)

p.mg.full <- ggrocs(rocs = list("Raw" = baseline_mg_roc, "Normalised" = full_mg_roc), 
        data_name = "Predicting coronoary artery disease", legend_title = "Manual Gating", 
        legend_order = c("Raw", "Normalised"), color_scale = my_colors_1, which_plot = "roc")

pdf(file = "../Plots/Figure_2_manual_gating_roc.pdf",   # The directory you want to save the file in
    width = 4, # The width of the plot in inches
    height = 4) # The height of the plot in inches
    p.mg.full
dev.off()

```