---
title: "Supplementary Figure 1"
author: "Elijah Willie"
date: "2025-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tidyverse)
  library(FuseSOM)
  library(ggsci)
  library(uwot)
  library(data.table)
  library(SingleCellExperiment)
  library(scales)
  library(ggnewscale)
  library(Polychrome)
  library(caret)
  library(RColorBrewer)
})
source('../scripts/normalisation_plotting.R')
set_random_seed(1994)
```

# load both studies
```{r}
nCores <- 4
BPPARAM <- simpleSeg:::generateBPParam(nCores)
theme_set(theme_classic())
```

# Set the markers
```{r}
useMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE')

allMarkers <- c('HLA_DR', 'CD3', 'CD4', 'CD8a', 'CD25', 'CD127', 'FoxP3', 'CD27',
                'KLRG1', 'CD56', 'CD45RO', 'CD45RA', 'CD192_CCR2', 'CD194_CCR4',
                'CD196_CCR6',
                'CD39', 'CD38', 'Ki67', 'CD183_CXCR3', 'CCR7', 'CD19', 'CD20',
                'IgD', 'CD14', 'CD304', 'CD141', 'CD1c_PE', "CD11b", "CD253_TRAIL",
                "CD34", "CD61", "CD11c", "eNOS", "LOX_1", "CD86", "CD16", "CD45_106",
                "CD45_108", "P2X7", "NOX5")
```

# Load in sce objects
```{r}
load('../../sce_dat/study_4_MMD_VAE_final.RData')
load('../../sce_dat/study_3_MMD_VAE_final.RData')
load('../../sce_dat/bioheart_cyCombine_batch_corrected.RData')
load('../../sce_dat/bioheart_iMUBAC_batch_corrected.RData')
```

# Update the markers
```{r}
useMarkers <- gsub("_", "-", useMarkers)
allMarkers <- gsub("_", "-", allMarkers)
```

# Define cell types of interest
```{r}
# only use celltypes that is present in both batches
final_celltypes <- c(
   "B cells",
   "CD8hi",
   "14+ monos",
   "CD4+ Tconv",
   "CD8lo",
   "NK",
   "16+ monos",
   "CD4+ Treg",
   "pDCs",
   "CD141+ DCs",
   "CD1c+ DCs"
    )
```


# Load in teh raw data for the BioHEART datasets
```{r}
df_4_raw <- assay(sce_norm, 'raw') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_norm$mg_cell_type_distinct) %>%
    dplyr::mutate(batch = sce_norm$CyTOF.Batch) %>%
    dplyr::mutate(sample_id = sce_norm$sample_id) %>%
    dplyr::mutate(Age = sce_norm$Age) %>%
    dplyr::mutate(Gender = sce_norm$Gender) %>%
    dplyr::mutate(Statin = sce_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_norm$gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_3_raw <- assay(sce_3_norm, 'raw') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_3_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_3_norm$CellTypes) %>%
    dplyr::mutate(batch = sce_3_norm$Batch) %>%
    dplyr::mutate(sample_id = sce_3_norm$sample_id) %>%
    dplyr::mutate(Age = sce_3_norm$Age) %>%
    dplyr::mutate(Gender = sce_3_norm$Gender) %>%
    dplyr::mutate(Statin = sce_3_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_3_norm$Gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_raw <- rbind(df_3_raw, df_4_raw)

```


# Load in the dioscRi normalised data for the BioHEART datasets
```{r}
df_4_norm <- assay(sce_norm, 'norm') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_norm$mg_cell_type_distinct) %>%
    dplyr::mutate(batch = sce_norm$CyTOF.Batch) %>%
    dplyr::mutate(sample_id = sce_norm$sample_id) %>%
    dplyr::mutate(Age = sce_norm$Age) %>%
    dplyr::mutate(Gender = sce_norm$Gender) %>%
    dplyr::mutate(Statin = sce_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_norm$gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_3_norm <- assay(sce_3_norm, 'norm') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype_new = sce_3_norm$Celltype_new) %>%
    dplyr::mutate(Celltype = sce_3_norm$CellTypes) %>%
    dplyr::mutate(batch = sce_3_norm$Batch) %>%
    dplyr::mutate(sample_id = sce_3_norm$sample_id) %>%
    dplyr::mutate(Age = sce_3_norm$Age) %>%
    dplyr::mutate(Gender = sce_3_norm$Gender) %>%
    dplyr::mutate(Statin = sce_3_norm$Statin) %>%
    dplyr::mutate(Gensini = as.factor(sce_3_norm$Gensini_bin)) %>%
    dplyr::filter(Celltype %in% final_celltypes)

df_norm <- rbind(df_3_norm, df_4_norm)
```

# Load in the cyCombine normalised data for the BioHEART datasets
```{r}
df_norm_cycombine <- assay(sce_cycombine, 'corrected') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype = sce_cycombine$CellType) %>%
    dplyr::mutate(sample_id = sce_cycombine$sample_id) %>%
    dplyr::mutate(batch = sce_cycombine$batch) %>%
    dplyr::mutate(Gensini = sce_cycombine$Gensini_bin) %>%
    dplyr::mutate(Statin = sce_cycombine$Statin) %>%
    dplyr::mutate(Age = sce_cycombine$Age) %>%
    dplyr::mutate(Gender = sce_cycombine$Gender) %>%
    # dplyr::filter(batch == 4)  %>%
    dplyr::filter(Celltype %in% final_celltypes)

```

# Load in the iMUBAC normalised data for the BioHEART datasets
```{r}
df_norm_imubac <- assay(sce_imubac, 'normexprs') %>%
  t() %>%
  as.data.frame() %>%
    dplyr::mutate(Celltype = sce_imubac$CellType) %>%
    dplyr::mutate(sample_id = sce_imubac$sample_id) %>%
    dplyr::mutate(batch = sce_imubac$batch) %>%
    dplyr::mutate(Gensini = sce_imubac$Gensini_bin) %>%
    dplyr::mutate(Statin = sce_imubac$Statin) %>%
    dplyr::mutate(Age = sce_imubac$Age) %>%
    dplyr::mutate(Gender = sce_imubac$Gender) %>%
    # dplyr::filter(batch == 4) %>%
    dplyr::filter(Celltype %in% final_celltypes)
```

```{r}
# put the data on the same scale as the normalised for comparison
preProcValues_raw <- preProcess(df_4_raw[, useMarkers], method = c("range"))
df_4_raw[, useMarkers] <- predict(preProcValues_raw, df_4_raw[, useMarkers])

preProcValues_cycombine <- preProcess(df_norm_cycombine[, useMarkers], method = c("range"))
df_norm_cycombine[, useMarkers] <- predict(preProcValues_cycombine, df_norm_cycombine[, useMarkers])

preProcValues_imubac <- preProcess(df_norm_imubac[, useMarkers], method = c("range"))
df_norm_imubac[, useMarkers] <- predict(preProcValues_imubac, df_norm_imubac[, useMarkers])

preProcValues_raw <- preProcess(df_raw[, useMarkers], method = c("range"))
df_raw[, useMarkers] <- predict(preProcValues_raw, df_raw[, useMarkers])
```



```{r}
samples <- unique(df_raw$sample_id)

t_raw <- get_summaries_per_channel(data = df_raw, 
                                   markers = useMarkers, samples = samples)

t_norm <- get_summaries_per_channel(data = df_norm, 
                                   markers = useMarkers, samples = samples)

t_norm_cycombine <- get_summaries_per_channel(data = df_norm_cycombine, 
                                   markers = useMarkers, samples = samples)

t_norm_imubac <- get_summaries_per_channel(data = df_norm_imubac, 
                                   markers = useMarkers, samples = samples)
```

## Supp Figure 1A
```{r}
p.box.sample <- boxSummaryValsPrePost_ggplot(raw = t_raw, dioscri = t_norm, 
                                         cycombine = t_norm_cycombine,
                                         imubac = t_norm_imubac, legendPos = "bottom",
                                         type = "both", title = "Mean of cell sample expression")

pdf(file = "../Plots/Supp_Figure_1_sample_box.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.box.sample
dev.off()
```

## Supp Figure 1B
```{r}
p.var.sample <- varBarPlot_ggplot(raw = t_raw, dioscri = t_norm, 
                                         cycombine = t_norm_cycombine,
                                         imubac = t_norm_imubac, legendPos = "bottom",
                                         type = "both", title = "Variance of sample expression")

pdf(file = "../Plots/Supp_Figure_1_sample_var_plot.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.var.sample
dev.off()
```

```{r}
celltypes <- final_celltypes

t_raw_mg <- get_summaries_per_celltype(data = df_raw, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_mg <- get_summaries_per_celltype(data = df_norm, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_cycombine_mg <- get_summaries_per_celltype(data = df_norm_cycombine, 
                                   markers = useMarkers, celltypes = celltypes)

t_norm_imubac_mg <- get_summaries_per_celltype(data = df_norm_imubac, 
                                   markers = useMarkers, celltypes = celltypes)
```

# Supp Figure 1C
```{r}
p.box.mg <- boxSummaryValsPrePost_ggplot(raw = t_raw_mg, dioscri = t_norm_mg, 
                                         cycombine = t_norm_cycombine_mg,
                                         imubac = t_norm_imubac_mg, legendPos = "bottom",
                                         type = "both", title = "Mean of cell type expression")

pdf(file = "../Plots/Supp_Figure_1_celltype_box.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.box.mg
dev.off()
```

# Supp Figure 1D
```{r}
p.var.mg <- varBarPlot_ggplot(raw = t_raw_mg, dioscri = t_norm_mg, 
                                         cycombine = t_norm_cycombine_mg,
                                         imubac = t_norm_imubac_mg, legendPos = "bottom",
                                         type = "both", title = "Variance of cell type expression")

pdf(file = "../Plots/Supp_Figure_1_celltype_var_plot.pdf",   # The directory you want to save the file in
    width = 6, # The width of the plot in inches
    height = 5) # The height of the plot in inches
    p.var.mg
dev.off()
```


