---
title: "Breast Cancer Prediction"
author: "Elijah WIllie"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
gc()
suppressPackageStartupMessages({
  library(tidyverse)
  library(FuseSOM)
  library(data.table)
  library(SingleCellExperiment)
  library(dioscRi)
})
```

# Load in data and compute the reference sample
```{r}
# Epithelial cell markers
useMarkers <- c("AKT","AR","BCL2","CA9", "CD24","CD44","CD49f", "cMET","cMYC",
                 "CyclinB1","ECadherin","EGFR","EpCAM","ERa","EZH2","H3K27me3", "HER2",
                "HLADR", "K14","K5","K7","K8K18","Ki67","p53","panK",
                "PRB","PTEN","SMA", "Survivin", "Vimentin"
                 )           

metadata <- read.csv('~/Documents/PhD/DeepLearning_CyTOF/Breast_Cancer_Wagner_2019/data/breast_cancer_metadata.csv')

metadata$dataset <- if_else(metadata$dataset == 'test', 'test', 'train')

```

# Load normalised data
```{r}
load('~/Documents/PhD/DeepLearning_CyTOF/Breast_Cancer_Wagner_2019/data/breast_cancer_train_MMD_VAE_sce.RData')
load('~/Documents/PhD/DeepLearning_CyTOF/Breast_Cancer_Wagner_2019/data/breast_cancer_test_MMD_VAE_sce.RData')
```

## cluster training data
```{r, eval=T}
# run FuseSOM clustering
nclust = 11
sce_norm <- runFuseSOM(sce_norm, numClusters = nclust, assay = 'norm',
                       verbose = F)
```

# Train clustering model with Caret
## Select training and testing data
```{r}
train_x <- reducedDim(sce_norm, type = "VAE") %>%
  mutate(cellTypes = sce_norm$clusters)

test_x <- reducedDim(sce_norm_test, type = "VAE")
```

```{r, eval=T}
test_clusters <- trainCellTypeClassifier(trainX = train_x, testX = test_x, model = 'lda')

table(test_clusters)

sce_norm_test$clusters <- test_clusters
```
# Set up clinical data
## Training
```{r}
clinicaldata <- colData(sce_norm) %>%
  as.data.frame() %>%
  dplyr::select(file_name, condition) %>%
  distinct() %>%
  dplyr::mutate(sample_id = file_name)
clinicaldata$sample_id <- as.character(clinicaldata$sample_id)

clinicaldata$condition <- factor(if_else(clinicaldata$condition == "T", 1, 0))
sce_norm$sample_id <- sce_norm$file_name
```

## Testing
```{r}
clinicaldata_test <- colData(sce_norm_test) %>%
  as.data.frame() %>%
  dplyr::select(file_name, condition) %>%
  distinct() %>%
  dplyr::mutate(sample_id = file_name)
clinicaldata_test$sample_id <- as.character(clinicaldata_test$sample_id)

clinicaldata_test$condition <- factor(if_else(clinicaldata_test$condition == "T", 1, 0))
sce_norm_test$sample_id <- sce_norm_test$file_name
```


# Compute training features
## Proportions
```{r}
data_logit <- computeFeatures(sce = sce_norm, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
row_names <- rownames(data_logit)
condition <- factor(clinicaldata[match(row_names,clinicaldata$sample_id),
                                      "condition"])
```

## Means
```{r, eval=T}
markerMeanCellType <- computeFeatures(sce = sce_norm, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```

## Compute features on test set
## Proportions
```{r}
data_test_logit <- computeFeatures(sce = sce_norm_test, featureType = 'prop', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)

row_names_test <- rownames(data_test_logit)
condition_test <- factor(clinicaldata_test[match(row_names_test,clinicaldata_test$sample_id),
                                      "condition"])
```

## Means
```{r, eval=T}
markerMeanCellType_test <- computeFeatures(sce = sce_norm_test, featureType = 'mean', 
                              cellTypeCol = 'clusters', sampleCol = 'sample_id', 
                              logit = T, useMarkers = useMarkers)
```


# Generate grouping structures
```{r}
nclust <- length(unique(test_clusters))
tree <- generateTree(features = data_logit, method = "ward")
tree <- tree$tree
groups <- generateGroups(tree = tree, nClust = nclust, 
                         proportions = data_logit, means = markerMeanCellType)
```

# Process training and testing data
## Training
```{r}
# Combine all input data matrices
X_train <- cbind(data_logit, markerMeanCellType) 
  
# Scale the combined data
scaleVals <- preProcess(X_train, method = c('scale'))
X_train <- predict(scaleVals, X_train) %>%
    as.matrix()
  
# Extract response variable
y_train <- as.numeric(levels(condition))[condition]
```

## Testing
```{r}
X_test <- cbind(data_test_logit, markerMeanCellType_test) %>%
  as.matrix()
X_test <- predict(scaleVals, X_test)
```


## Fit Overlap model
```{r}
fit <- fitModel(xTrain = X_train, yTrain = y_train,
                groups = groups, penalty = 'grLasso', seed = 1994)
```



## Predict on test data
```{r}
y_test <- as.numeric(levels(condition_test))[condition_test]
test_auc <- plotAUC(fit = fit$fit, xTest = X_test, yTest = y_test, title = "Test set AUC =")
test_auc$plot

# write.csv(test_auc$preds,
#           '../data/breast_cancer_dioscRi_aucs.csv')
```
# Visualise Resulting Tree
```{r, fig.height=10, fig.width=18}
visualiseModelTree(fit = fit$fit, tree = tree, type = "cluster", 
                   trainingData = X_train)
```
